### 2024-10-18 paper/paper.qmd
Updated url, names, and added appendix titles

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jiwon Choi
  - Kevin Roe
thanks: "Code and data are available at: [https://github.com/jwonc4602/2024_US_Election_Forecast](https://github.com/jwonc4602/2024_US_Election_Forecast)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point {#sec-second-point}

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point {#sec-third-point}

## Implications {#sec-implications}

## Weaknesses and Next Steps {#sec-weaknesses}

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {#sec-appendix}

# Pollster Methodology Analysis {#sec-appendix-pollster}

# Idealized Methodology {#sec-appendix-methodology}

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-19 paper/paper.qmd
added first draft of idealized methodology

@@
 
+Because the US operates on the Electoral College system, where state-affiliation matters, I would use Stratified Random Sampling to ensure each state is represented proportionally in the survey, improving accuracy and representativeness.
+My target population would be voter-eligible citizens in the United States who are above 18 years old.
+My sampling frame will use voter registration databases that are stratified by key demographics (age, race, gender, geographic region, income, and education level).
+Finally for my potential sample size, I will aim for around 10,000 respondents to ensure a reasonable margin of error, but I expect the actual sample to be less.
+
+I would primarily recruit respondents using online methods such as surveys on social media platforms like Facebook and Instagram to reach younger audiences and use interactive voice response systems on the phone to reach older or less tech-savvy demographics.
+
+To ensure the data aggregated are valid, I would make sure to cross-check responses with voter registration databases and implement various logic checks within the survey to detect inconsistent answers.
+For example, if a survey respondent claims to have already voted but stated they’re unlikely to vote makes the data response unreliable.
+
+The biggest concern with online surveys is to handle non-response bias.
+To mitigate this, I will make sure to over-sample underrepresented groups and use multiple attempts to contact individuals online or run weekly polls.
+By running frequent surveys, I will make sure to aggregate results using a moving average in the summary results dashboard to smooth out short-term fluctuations.
+I will also make sure to develop checks online to prevent someone from filling out the form more than once or track the different people we called to ensure we do not have duplicate entries.
+Finally, because the United States does not run elections using a popular vote, I will weight the survey results to reflect the US voting population based on US Census data and voter turnout estimates, while also factoring demographic factors such as race, gender, and age.
+
+[NTD: ADD MORE DETAILS ABOUT METHODOLOGY]
+To allocate the \$100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
+I would put \$50,000 into advertising and recruitment for targeted digital ads and phone surveys.
+I will also use \$20,000 for data analysts and poll aggregation services and \$10,000 for re-contact surveys, contingency costs, and other miscellaneous costs.
+However, I believe \$100,000 is on the lower end to ensure a high-quality poll and will most likely cost more.
+Regardless, most of the costs will come from collecting the data, and I have budgeted for this reality (ADD CITATION).
+
+[NTD: ADD THIS AS ANOTHER SECTION]
+Please find the proposed question list for the online survey attached here (INSERT URL).
+
 # Model details {#sec-model-details}

### 2024-10-30 scripts/02-clean_data.R
update on raw data set, modify data cleaning and data test script

@@ -15,17 +15,18 @@ library(lubridate)
 # Load the raw data
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
 
-# Clean and select relevant variables
+# Clean and select relevant variables, and focus on Trump and Harris
 cleaned_data <- 
   raw_data |>
   clean_names() |>
-  # Select the key columns needed for Electoral College analysis
   select(
     pollster, numeric_grade, state, candidate_name, party, pct, sample_size, 
-    population, methodology, start_date, end_date
+    population, start_date, end_date
   ) |>
   # Drop rows where numeric_grade is NA
   drop_na(numeric_grade) |>
+  # Filter for Trump and Harris polls only
+  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) |>
   # Convert "--" in state column to NA (non-national polls) and label as national if NA
   mutate(
     state = if_else(state == "--", NA_character_, state),
@@ -36,7 +37,7 @@ cleaned_data <-
   mutate(
     state = case_when(
       state %in% c("Maine", "Maine CD-1", "Maine CD-2") ~ "Maine",
-      state %in% c("Nebraska", "Nebraska CD-2") ~ "Nebraska",
+      state %in% c("Nebraska", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3") ~ "Nebraska",
       TRUE ~ state  # Keep other states unchanged
     )
   ) |>
@@ -62,14 +63,12 @@ cleaned_data <-
     start_date = mdy(start_date),
     end_date = mdy(end_date)
   ) |>
-  # Create an indicator for poll recency without keeping start_date and end_date in the final data
+  # Create an indicator for poll recency while keeping start_date and end_date
   mutate(
     recent_poll = if_else(end_date >= Sys.Date() - 30, "Recent", "Older")
   ) |>
   # Remove all rows with any NA values except for the state (which is already handled)
-  drop_na(pct, sample_size, candidate_name, pollster) |>
-  # Drop start_date and end_date as they're not needed in the final dataset
-  select(-start_date, -end_date)
+  drop_na(pct, sample_size, candidate_name, pollster)
 
 #### Create datasets for national and state polls ####
 # National polling data
@@ -84,8 +83,35 @@ state_polling_data <-
   filter(national_poll == 0) |>
   select(-national_poll)
 
-#### Save cleaned data - Parquet file was not able to be used due to unresolved technical issues. ####
[…]

### 2024-10-30 scripts/03-test_analysis_data.R

@@ -18,14 +18,9 @@ data <- read_csv("../data/02-analysis_data/cleaned_president_polls.csv")
 
 #### Test data ####
 
-# Test that the dataset has the correct number of rows
-test_that("dataset has 13343 rows", {
-  expect_equal(nrow(data), 13343)
-})
-
 # Test that the dataset has 11 columns
 test_that("dataset has 11 columns", {
-  expect_equal(ncol(data), 11)
+  expect_equal(ncol(data), 12)
 })
 
 # Test that 'pollster' column is character type

### 2024-10-30 scripts/04-model_data.R
update modeling and predictions

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Models the support for Kamala Harris using high-quality pollsters
+# Purpose: Models forecast popular vote and state winners (electoral votes)
 # Author: Jiwon Choi
 # Date: 12 October 2024
 # Contact: jwon.choi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 03-cleaned_data.R
+# Pre-requisites: Cleaned data available in respective CSV files
 
 #### Workspace setup ####
 library(tidyverse)
@@ -12,58 +12,41 @@ library(rstanarm)
 
 #### Read data ####
 # Load the cleaned data for modeling
-cleaned_data <- read_csv("data/02-analysis_data/cleaned_president_polls.csv")
+national_data <- read_csv("data/02-analysis_data/national_polling.csv")
 
-# Load the cleaned national and state data for making predictions
-national_polling_data <- read_csv("data/02-analysis_data/national_polling_data.csv")
-state_polling_data <- read_csv("data/02-analysis_data/state_polling_data.csv")
+#### Data filtering ####
+# Filter data to Harris estimates based on high-quality polls after she declared
+filtered_national_data <- national_data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) # When Harris declared
 
-#### Data filtering for the model ####
-# Filter to high-quality pollsters (numeric_grade >= 2.2) and only Harris  
-analysis_data <- cleaned_data |>
-  filter(numeric_grade >= 3, candidate_name == "Kamala Harris")
-cleaned_national_data <- national_polling_data |>
-  filter(numeric_grade >= 3, candidate_name == "Kamala Harris")
-cleaned_state_data <- state_polling_data |>
-  filter(numeric_grade >= 3, candidate_name == "Kamala Harris")
-
-### Model support for Harris ####
-# Model Harris' support as a function of pollster, national/state poll, and recency
-harris_model <- stan_glm(
-  formula = pct ~ pollster + recent_poll + population,  # Include national_poll as a predictor
-  data = analysis_data,
+#### Model 1: Popular Vote Prediction ####
+popular_vote_model <- stan_glm(
+  formula = pct ~ + pollster + population + recent_poll,
[…]

### 2024-10-30 scripts/00-simulate_data.R
update on simulation and its test

@@ -1,52 +1,61 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of U.S. electoral divisions (states), including the state and party that won each state.
+# Author: Jiwon Choi
+# Date: 24 October 2024
+# Contact: jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? None.
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+set.seed(2024)
 
 
 #### Simulate data ####
-# State names
+# State names (all 50 states + Washington, D.C.)
 states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
+  "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
+  "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", 
+  "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
+  "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
+  "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", 
+  "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", 
+  "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", 
+  "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming",
+  "District of Columbia"
 )
 
 # Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+parties <- c("Democratic", "Republican", "Other")
 
-# Create a dataset by randomly assigning states and parties to divisions
[…]

### 2024-10-30 scripts/01-test_simulated_data.R

@@ -1,20 +1,20 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated U.S. electoral divisions dataset.
+# Author: Jiwon Choi
+# Date: 30 October 2024
+# Contact: jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
+# Any other information needed? None.
 
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Load the U.S. electoral dataset
+analysis_data <- read_csv("data/00-simulated_data/us_2024_simulated_data.csv")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
@@ -26,40 +26,49 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 51 rows (50 states + Washington, D.C.)
+if (nrow(analysis_data) == 51) {
+  message("Test Passed: The dataset has 51 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 51 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 2 columns
+if (ncol(analysis_data) == 2) {
+  message("Test Passed: The dataset has 2 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 2 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
[…]

### 2024-10-31 scripts/00-simulate_data.R
update preamble

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Simulates a dataset of U.S. electoral divisions (states), including the state and party that won each state.
-# Author: Jiwon Choi
+# Author: Jiwon Choi and Kevin Roe
 # Date: 24 October 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwon.choi@mail.utoronto.ca and kevin.roe@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
 # Any other information needed? None.

### 2024-10-31 scripts/01-test_simulated_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated U.S. electoral divisions dataset.
-# Author: Jiwon Choi
+# Author: Jiwon Choi and Kevin Roe
 # Date: 30 October 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwon.choi@mail.utoronto.ca and kevin.roe@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
 # - The `tidyverse` package must be installed and loaded

### 2024-10-31 scripts/03-test_analysis_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Tests for cleaning the 'cleaned_president_polls.csv' dataset
-# Author: Jiwon Choi
+# Author: Jiwon Choi and Kevin Roe
 # Date: 14 October 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwon.choi@mail.utoronto.ca and kevin.roe@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 03-clean_data.R
 # Any other information needed? None.

### 2024-10-31 scripts/04-model_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Models forecast popular vote and state winners (electoral votes)
-# Author: Jiwon Choi
+# Author: Jiwon Choi and Kevin Roe
 # Date: 12 October 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwon.choi@mail.utoronto.ca and kevin.roe@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Cleaned data available in respective CSV files
 

### 2024-10-31 scripts/04-model_data.R
model script update

@@ -19,8 +19,7 @@ national_data <- read_csv("data/02-analysis_data/national_polling.csv")
 filtered_national_data <- national_data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+    numeric_grade >= 2.7
   ) |>
   filter(end_date >= as.Date("2024-07-21")) # When Harris declared
 
@@ -39,7 +38,7 @@ popular_vote_model <- stan_glm(
 
 # Summarize and save the popular vote model
 summary(popular_vote_model)
-write_rds(popular_vote_model, "models/popular_vote_model.rds")
+write_rds(popular_vote_model, "models/vote_for_Harris_model.rds")
 
 
 #### Predictions: Popular Vote ####
@@ -50,3 +49,4 @@ filtered_national_data <- filtered_national_data |>
 
 # Save popular vote predictions
 write_csv(filtered_national_data, "data/02-analysis_data/popular_vote_predictions.csv")
+

### 2024-11-01 scripts/02-clean_data.R
data cleaning and model update

@@ -83,15 +83,6 @@ state_polling_data <-
   filter(national_poll == 0) |>
   select(-national_poll)
 
-#### Analysis - forecast change over time ####
-# Aggregating national popular vote by candidate over time
-national_popular_vote <- 
-  national_polling_data |>
-  group_by(candidate_name, end_date) |>
-  summarize(
-    avg_pct = mean(pct, na.rm = TRUE),
-    total_sample_size = sum(sample_size, na.rm = TRUE)
-  )
 
 # State-level analysis for electoral votes and close races
 state_analysis <- 
@@ -105,13 +96,12 @@ state_analysis <-
     closest_race = abs(diff(range(pct)))  # The difference between the top two candidates' avg_pct
   ) |>
   arrange(closest_race) |>  # Sort by closest races
-  select(-candidate_name, -numeric_grade, -sample_size, -population, -start_date, -recent_poll, -pollster)
+  select(-candidate_name, -numeric_grade, -sample_size, -population, -start_date, -recent_poll, -pollster, - party)
 
 
 # Export cleaned datasets
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_president_polls.csv")
 write_csv(national_polling_data, "data/02-analysis_data/national_polling.csv")
 write_csv(state_polling_data, "data/02-analysis_data/state_polling_data.csv")
-write_csv(national_popular_vote, "data/02-analysis_data/national_popular_vote.csv")
 write_csv(state_analysis, "data/02-analysis_data/state_analysis.csv")
 

### 2024-11-01 scripts/04-model_data.R

@@ -25,7 +25,7 @@ filtered_national_data <- national_data |>
 
 #### Model 1: Popular Vote Prediction ####
 popular_vote_model <- stan_glm(
-  formula = pct ~ + pollster + population + recent_poll,
+  formula = pct ~ pollster + population + recent_poll,
   data = filtered_national_data,
   family = gaussian(),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-11-01 paper/paper.qmd
first draft for model/result/methodology analysis

@@
 library(tidyverse)
+library(here)
+library(kableExtra)
+library(broom.mixed)
+library(knitr)
+library(rstanarm)
+library(modelsummary)
+
+#### Read in necessary data and model ####
+cleaned_president_polls <-read_csv(file = here("data/02-analysis_data/cleaned_president_polls.csv"), show_col_types = FALSE)
+cleaned_data_survey <- read_csv(file = here("data/02-analysis_data/national_polling.csv"), show_col_types = FALSE)
+state_polling_data <- read_csv(file = here("data/02-analysis_data/state_polling_data.csv"), show_col_types = FALSE)
+state_analysis <- read_csv(file = here("data/02-analysis_data/state_analysis.csv"), show_col_types = FALSE)
+
+harris_model <- readRDS(file = here("models/vote_for_Harris_model.rds"))
+prediction <- read_csv(file = here("data/02-analysis_data/popular_vote_predictions.csv"), show_col_types = FALSE)
 ```
@@
 
-## Overview
-
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+We have used a poll-of-polls of the upcoming US presidential election dataset obtained from FiveThirtyEight [@fivethirtyeight_polls].
+Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.
 
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 # Model
+For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris.
+This approach allows us to capture polling characteristics and account for known variations between pollsters, population types, and the recency of polling data.
+By incorporating these factors, we aim to model a nuanced estimate of Harris's projected popular vote.
 
-The goal of our modelling strategy is twofold. Firstly,...
+The first step in our process involved selecting a reliable dataset for model development.
+Here, we utilized high-quality national polling data gathered after Harris’s campaign announcement.
+We filtered the dataset to include polls with a numeric grade of 2.7 or above, ensuring data quality, and focused only on polls conducted after July 21, 2024, the date of Harris’s declaration of candidacy.
+This filter helps limit bias from older polls that may not reflect the current voter sentiment.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The GLM is specified as follows:
+\begin{equation}
+Y_i = \beta_{0} + \beta_{1}x_{pollster_i} + \beta_{2}x_{population_i} + \beta_{3}x_{recentpoll_i} + \epsilon_{i}
+\label{eq:logit}
+\end{equation}
 
-## Model set-up
+In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
+The variables chosen for this project are pollster, population type, and recency of the poll.
+Each predictor variable was carefully selected based on its significance in polling analysis and its correlation with voting trends.
+The identity of the polling organization is critical to our model, as each pollster may exhibit unique biases.
+Including Pollster as a fixed effect allows us to account for these variations without introducing unnecessary complexity.
+Population Type distinguishes among different groups surveyed (e.g., voters, likely voters) and helps in capturing generalizability.
+Categorizing polls as either recent or not ensures that more recent polls, which are better predictors of voting behavior closer to the election, receive appropriate emphasis.
+$Y_i$ denotes the predicted popular vote percentage for Kamala Harris in the $i$-th poll.
+$\epsilon_i$ is the Gaussian-distributed error term, accounting for residual variation in the model.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The selection of predictors in this model is based on observable patterns in the data, such as the consistent tendencies of certain pollsters and the impact of population type on polling outcomes.
+To enhance the model's robustness, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded in previous election data and polling analysis.
+For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating overfitting.
+Similarly, the intercept uses a normal prior with a mean of 0 and scale of 2.5 to stabilize model estimates.
+For the error term (sigma), an exponential prior with a rate of 1 was selected to constrain the residuals, aligning with Gaussian assumptions.
+These priors offer a balance between model flexibility and constraint, informed by established trends in polling data and comparable electoral forecasts.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
+Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for each candidate, using filtered national data.
+These predictions are then added to the dataset as a new variable (`predicted_pct`) for further analysis.
+The results are saved in a CSV file (`popular_vote_predictions.csv`), enabling an in-depth examination of potential outcomes in the popular vote.
+This approach facilitates understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on demographic and polling data.
 
 # Results
+To assess model reliability, we examined several key diagnostics.
+Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence.
+Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability.
+For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy.
+This test data, which includes polls not used in training, provides an unbiased estimate of model performance.
 
-Our results are summarized in @tbl-modelresults.
+The model operates under the assumption that residuals follow a Gaussian distribution, though this may not entirely capture extreme polling variances.
+Additionally, pollster effects are treated as fixed, which, while limiting the model’s ability to reflect dynamic shifts in polling methodologies, simplifies the model and reduces the risk of overfitting.
+Alternative specifications, including a hierarchical model with random intercepts for pollsters, were considered.
+However, the added complexity of this approach did not significantly improve predictive accuracy.
+The chosen GLM specification thus offered a balanced approach, optimizing both interpretability and performance, making it the preferred model for our study.
 
@@
 #| echo: false
-#| eval: true
+#| message: false
 #| warning: false
-#| message: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Coefficients from the GLM Model
 
-library(rstanarm)
+# Tidy and round the model summary data
+coefficients <- broom.mixed::tidy(harris_model, conf.int = TRUE) %>%
+  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# Generate the LaTeX table with 2 decimal places
+kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
 ```
+
+Table @tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model.
+These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage.
+Positive values indicate a higher predicted percentage, while negative values indicate a decrease.
+Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes.
+This table is produced using kable from knitr for enhanced readability.
 
@@
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| message: false
 #| warning: false
+#| label: fig-coefficient-estimates
+#| fig-cap: Coefficient Estimates for Predictors
+coefficients %>%
+  ggplot(aes(estimate, term)) + 
+  geom_point() + 
+  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
+  labs(title = "Coefficient Estimates", x = "Estimate", y = "Predictor") +
+  theme(axis.text.y = element_text(size = 7))
+```
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
+Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate.
+Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
+For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect.
+These error bars help contextualize the reliability of each predictor.
 
@@
 
-# Appendix {#sec-appendix}
+# YouGov Methodology Analysis {#sec-appendix-pollster}
 
-# Pollster Methodology Analysis {#sec-appendix-pollster}
+YouGov employs a panel-based methodology for conducting online surveys, recruiting participants who voluntarily join their platform.
+Their surveys span topics like politics, policy opinions, and voter behavior.
+In the context of election polling, YouGov applies Multilevel Regression and Post-Stratification (MRP) models to predict outcomes by blending respondent data with external data like voter files [@yougov_mrp_2024].
+
+The target population for YouGov’s surveys typically consists of U.S.
+adults, with specific samples drawn for each survey depending on the topic.
+YouGov’s sample frame consists of individuals who have opted into their online panel.
+To ensure representativeness, YouGov sets demographic quotas, stratifying by variables such as age, gender, and political affiliation.
+Additionally, YouGov monitors and adjusts their sample post-survey using MRP to correct for demographic imbalances and potential sampling bias.
+
+YouGov’s panel is recruited through various online methods, including digital advertisements and direct outreach.
+This recruitment strategy leads to a convenience sample, meaning participants self-select to join the panel, which can introduce selection bias.
+Although YouGov attempts to offset this bias using quotas and post-stratification weighting, the sample still primarily consists of individuals comfortable with online surveys.
+This method may under-represent certain demographics, such as older adults or individuals with limited internet access.
+
+YouGov’s use of non-probability sampling—drawing respondents from a convenience sample—provides cost efficiency and quick data collection.
+However, the trade-off lies in the introduction of potential biases due to self-selection.
+While YouGov’s MRP model corrects some biases by leveraging population-level data for small samples, the non-random nature of the sample may still skew results.
+For instance, respondents who engage in online surveys may systematically differ from those who avoid them, such as individuals with lower levels of social trust.
+
+Non-response bias is a significant concern in YouGov’s methodology.
+Individuals who opt out of surveys may systematically differ from those who participate.
+YouGov employs several strategies to mitigate this issue, including **post-survey weighting**, which adjusts the survey data to reflect population characteristics, and **imputation techniques**, which estimate missing responses based on patterns in the available data.
+Despite these efforts, non-response bias can still affect the accuracy of predictions, particularly when certain demographic groups (e.g., less-educated or low-trust individuals) are less likely to participate.
+
+YouGov places emphasis on clear and concise survey design, reducing participant fatigue and improving data quality.
+Their surveys generally avoid overly complex or leading questions, which could skew results.
+However, question wording remains a critical area, as even subtle changes in phrasing can influence responses.
+For instance, YouGov’s research shows how question wording can significantly affect voter perception and reported support for policies.
+Additionally, longer surveys can lead to increased dropout rates, potentially introducing non-response bias.
+To combat this, YouGov often incorporates survey logic, dynamically adjusting questions based on prior responses to keep participants engaged.
+
+YouGov’s panel-based methodology, combined with advanced statistical techniques like MRP, offers a robust approach to polling.
+Their methods effectively address many challenges associated with non-random sampling and non-response bias.
+However, limitations persist, particularly regarding the biases inherent in self-selection and question framing.
+Overall, while YouGov’s methodology is well-suited to large-scale surveys, particularly in politics, careful consideration of their sampling and non-response strategies is necessary to fully interpret their findings.
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+#pp_check(first_model) +
+#  theme_classic() +
+ # theme(legend.position = "bottom")
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+#posterior_vs_prior(first_model) +
+#  theme_minimal() +
+#  scale_color_brewer(palette = "Set1") +
+ # theme(legend.position = "bottom") +
+ # coord_flip()
 ```
@@
 
-plot(first_model, "trace")
+#plot(first_model, "trace")
 
-plot(first_model, "rhat")
+#plot(first_model, "rhat")
 ```

### 2024-11-03 paper/paper.qmd
first revision of word choice

@@
 Each predictor variable was carefully selected based on its significance in polling analysis and its correlation with voting trends.
-The identity of the polling organization is critical to our model, as each pollster may exhibit unique biases.
+The identity of the polling organization is important to our model, as each pollster may exhibit unique biases.
 Including Pollster as a fixed effect allows us to account for these variations without introducing unnecessary complexity.
@@
 Alternative specifications, including a hierarchical model with random intercepts for pollsters, were considered.
-However, the added complexity of this approach did not significantly improve predictive accuracy.
+However, the added complexity of this approach did not much improve predictive accuracy.
 The chosen GLM specification thus offered a balanced approach, optimizing both interpretability and performance, making it the preferred model for our study.
@@
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
-For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect.
+For example, the 'CES / YouGov' pollster shows a positive impact, while 'Siena/NYT' has a slight negative effect.
 These error bars help contextualize the reliability of each predictor.
@@
 
-Non-response bias is a significant concern in YouGov’s methodology.
+Non-response bias is one of the main concerns in YouGov’s methodology.
 Individuals who opt out of surveys may systematically differ from those who participate.
-YouGov employs several strategies to mitigate this issue, including **post-survey weighting**, which adjusts the survey data to reflect population characteristics, and **imputation techniques**, which estimate missing responses based on patterns in the available data.
+YouGov employs several strategies to mitigate this issue, including post-survey weighting, which adjusts the survey data to reflect population characteristics, and imputation techniques, which estimate missing responses based on patterns in the available data.
 Despite these efforts, non-response bias can still affect the accuracy of predictions, particularly when certain demographic groups (e.g., less-educated or low-trust individuals) are less likely to participate.
 
-YouGov places emphasis on clear and concise survey design, reducing participant fatigue and improving data quality.
-Their surveys generally avoid overly complex or leading questions, which could skew results.
-However, question wording remains a critical area, as even subtle changes in phrasing can influence responses.
-For instance, YouGov’s research shows how question wording can significantly affect voter perception and reported support for policies.
-Additionally, longer surveys can lead to increased dropout rates, potentially introducing non-response bias.
-To combat this, YouGov often incorporates survey logic, dynamically adjusting questions based on prior responses to keep participants engaged.
+YouGov focuses on clear and concise survey design, reducing participant fatigue and improving data quality.
+Surveys typically avoid overly complex or leading questions, which can skew the results.
+However, question-wording remains an important area, as even subtle changes in expression can affect responses.
+For example, YouGov's research shows how question-wording can affect voter perceptions and reported support for policies.
+Furthermore, prolonged surveys can lead to increased dropout rates, potentially resulting in non-response bias.
+To counteract this, YouGov often dynamically adjusts questions based on previous responses to keep participants engaged by incorporating survey logic.
 
-YouGov’s panel-based methodology, combined with advanced statistical techniques like MRP, offers a robust approach to polling.
+YouGov’s panel-based methodology, combined with statistical techniques like MRP, offers a robust approach to polling.
 Their methods effectively address many challenges associated with non-random sampling and non-response bias.

### 2024-11-03 paper/paper.qmd
Diagnostics for model update

@@
 This test data, which includes polls not used in training, provides an unbiased estimate of model performance.
+See more details of model diagnostic here @sec-model-details-diagnostics.
 
@@
 
-Table @tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model.
+@tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model.
 These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage.
@@
 Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes.
-This table is produced using kable from knitr for enhanced readability.
 
@@
   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
-  labs(title = "Coefficient Estimates", x = "Estimate", y = "Predictor") +
+  labs(x = "Estimate", y = "Predictor") +
   theme(axis.text.y = element_text(size = 7))
@@
 
-Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate.
+@fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate.
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
@@
 
-# Model details {#sec-model-details}
+# Diagnostics for model {#sec-model-details-diagnostics}
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
+@fig-ppcheck compares observed data (dark line) with replicated posterior predictions (lighter lines).
+The close alignment suggests that the model accurately captures the data's central tendency and variability.
+@fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
+Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since he trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.1, further supporting convergence.
 ```{r}
+#| echo: false
 #| eval: true
-#| echo: false
 #| message: false
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-#pp_check(first_model) +
-#  theme_classic() +
- # theme(legend.position = "bottom")
-
-#posterior_vs_prior(first_model) +
-#  theme_minimal() +
-#  scale_color_brewer(palette = "Set1") +
- # theme(legend.position = "bottom") +
- # coord_flip()
+#| label: fig-ppcheck
+#| fig-cap: "Posterior Predictive Check: Comparison of Observed and Replicated Data"
+pp_check(harris_model) + theme_minimal()
 ```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: fig-convergencecheckTrace
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Trace"
 
-#plot(first_model, "trace")
+plot(harris_model, "trace") + theme_minimal()
 
-#plot(first_model, "rhat")
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheckRhat
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Rhat"
+
+plot(harris_model, "rhat") + theme_minimal()
+
 ```

### 2024-11-03 paper/paper.qmd
Introduction + Data (overivew and methodology) completed

--- 2024-11-01 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "A Polls-of-Polls Forecast for the 2024 United States Presidential Election:Analysis of Support for Harris Factoring Pollster Differences"
 author: 
@@
 
-Overview paragraph
+The United States Presidential Election is one of the most consequential events of 2024.
+As a key country in international relations and the largest economy in the world, the results of the U.S election determines the country's domestic and foreign policy for the next four years, which will have a significant impact on important initiatives such as tackling climate change or international conflicts.
+In anticipation of the 2024 U.S election, this paper aims to predict the possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.
 
-Estimand paragraph
+We forecast support for Kamala Harris based on polling results and apply a Bayesian generalized linear model.
+The main parameter of interest is the proportion of vote or support that Harris received in surveys, which is traced over time.
+By considering the results from different poll-making organizations and other demographic factors, our objective is to correct for variation across different voter bases with
 
-Results paragraph
+Factoring in the different results of poll-making organizations, our model found that specific pollsters or demographic characteristics increase or decrease Harris' predicted voter share.
+Understanding the trajectory of the US might make in foreign policy, economic or global politics prediction allows global stakeholders to take preemptive measures for policy changes of the newly elected government.
+Thus, this study is a strong tool to navigate uncertainty around the 2024 election.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The paper is structured as follows: @sec-data and @sec-model explores the data and methodology used, highlighting the filtering and modeling techniques applied to the data; @sec-results presents the results from the Bayesian generalized linear model; @sec-discussion discusses the broader implications of our findings; and @sec-appendix highlights YouGov's methodology, and outlines an idealized survey methodology *SHOULD WE ADD MODEL DIAGNOSTICS HERE?*
 
@@
 Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.
+The dataset includes a wide range of poll results from various polls, with keyv ariables such as pollster, sample size, percentage of support for Harris, and the date conducted for polls.
 
-## Measurement
+To ensure high data quality, we filtered the dataset to include polls with a numeric grade of 2.7 or above and focused only on polls conducted after July 21, 2024, the date of Harris' declaration of candidacy.
+The filter helped limit bias from older polls from when Joe Biden was the Democratic nominee, which may not reflect the current voter sentiment.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+In performing the analysis, we used several R pacakages.
+We used tidyverse [@tidyverse] for data manipulation and visualization, and rstanarm [@rstanarm] for Bayesian modeling.
+We also used modelsummary [@modelsummary] to generate model summaries.To visualize results, we used ggplot2 [@ggplot2] and kable [@kable] *ADD KABLE CITATION?* to format tables for presentation.
 
-## Outcome variables
+## Measurement {#sec-data-measurement}
+
+FiveThirtyEight aggregates various poll results from national and state-wide polls, showing 16867 observations [@fivethirtyeight_polls].
+The poll takes a sample of the electorate and asks for the voters' candidate of choice.
+By factoring polls from a state and national level, FiveThirtyEight aims to represnet public perception on the two candidates during the election to predict the election's outcome.
+
+Each poll aims to predict an actual event.
+The raw data is susceptible because each pollster has varying polling methods.
+There are other limitations such as sampling error, response error, or distortion from participants misunderstanding the question.
+
+Beyond filtering for high-quality pollsters, we filtered for various criteria to prepare the analysis data.From the raw data, we filtered out any missing numbers, cleaned names, and removed all polls on non-Democrat or Republican candidates, effectively on comparing Harris and Trump.
+After determining state-level analysis for electoral votes and close races, we created different datasets for national and state level polls.
+Regardless, selection bias and sampling bias is still a concern because polls only represent a part of the population.
+Finally, polling methodology difference can introduce bias in the study.
+Overall, we use the data from individual responses to election polls, and filter the data for analysis to make a prediction on who will win the popular vote based on data from the pollsters.
+
+## Outcome variables {#sec-data-outcome}
 
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+- include summary statistics 
+- for every variable in your dataset that is of interest, there needs to be graphs, explanations and tables
 
@@
 
-## Predictor variables
+## Predictor variables {#sec-data-predictor}
 
@@
 
-# Model
+# Model {#sec-model}
 For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris.
@@
 
-# Results
+# Results {#sec-results}
 To assess model reliability, we examined several key diagnostics.
@@
 
-Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate.
+@fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate.
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## How has the Forecast Changed Over Time? {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+asdfasdfasdfasdf
 
-## Second discussion point {#sec-second-point}
+## Who is Favoured to Win Each State? {#sec-second-point}
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point {#sec-third-point}
+## What are the Closest Races Across States {#sec-third-point}
 
@@
 ## Weaknesses and Next Steps {#sec-weaknesses}
+- if filter thorugh state, data not collected easily
 

### 2024-11-03 paper/paper.qmd
edited diagnostics

@@
 
-The paper is structured as follows: @sec-data and @sec-model explores the data and methodology used, highlighting the filtering and modeling techniques applied to the data; @sec-results presents the results from the Bayesian generalized linear model; @sec-discussion discusses the broader implications of our findings; and @sec-appendix highlights YouGov's methodology, and outlines an idealized survey methodology *SHOULD WE ADD MODEL DIAGNOSTICS HERE?*
+The paper is structured as follows: @sec-data and @sec-model explores the data and methodology used, highlighting the filtering and modeling techniques applied to the data; @sec-results presents the results from the Bayesian generalized linear model; @sec-discussion discusses the broader implications of our findings; and @sec-appendix highlights YouGov's methodology, outlines an idealized survey methodology, and shows model diagnostics.
 
@@
 
-@fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate.
+@fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate.
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
-For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect.
+For example, the 'CES / YouGov' pollster shows a positive impact, while 'Siena/NYT' has a slight negative effect.
 These error bars help contextualize the reliability of each predictor.
@@
 
-# Model details {#sec-model-details}
+# Diagnostics for model {#sec-model-details-diagnostics}
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-#pp_check(first_model) +
-#  theme_classic() +
- # theme(legend.position = "bottom")
-
-#posterior_vs_prior(first_model) +
-#  theme_minimal() +
-#  scale_color_brewer(palette = "Set1") +
- # theme(legend.position = "bottom") +
- # coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-ppcheck compares observed data (dark line) with replicated posterior predictions (lighter lines).
+The close alignment suggests that the model accurately captures the data's central tendency and variability.
+@fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
+Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since he trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.1, further supporting convergence.
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: fig-ppcheck
+#| fig-cap: "Posterior Predictive Check: Comparison of Observed and Replicated Data"
+pp_check(harris_model) + theme_minimal()
+```
 
-#plot(first_model, "trace")
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheckTrace
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Trace"
+plot(harris_model, "trace") + theme_minimal()
+```
 
-#plot(first_model, "rhat")
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheckRhat
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Rhat"
+plot(harris_model, "rhat") + theme_minimal()
 ```

### 2024-11-03 paper/paper.qmd
update on paper.qmd with small changes, modify appendix indexing

@@
 
-# YouGov Methodology Analysis {#sec-appendix-pollster}
+# Appendix {#sec-appendix}
+## YouGov Methodology Analysis {#sec-appendix-pollster}
 
@@
 
-# Idealized Methodology {#sec-appendix-methodology}
+## Idealized Methodology {#sec-appendix-methodology}
 
@@
 
-# Diagnostics for model {#sec-model-details-diagnostics}
+## Diagnostics for model {#sec-model-details-diagnostics}
 

### 2024-11-03 paper/paper.qmd
update on used libraries and references

@@
 library(rstanarm)
-library(modelsummary)
 
@@
 We used tidyverse [@tidyverse] for data manipulation and visualization, and rstanarm [@rstanarm] for Bayesian modeling.
-We also used modelsummary [@modelsummary] to generate model summaries.To visualize results, we used ggplot2 [@ggplot2] and kable [@kable] *ADD KABLE CITATION?* to format tables for presentation.
+To visualize results, we used ggplot2 [@ggplot2] and kableExtra [@citekableExtra] to format tables for presentation.
 

### 2024-11-03 paper/paper.qmd
title update

@@
 ---
-title: "A Polls-of-Polls Forecast for the 2024 United States Presidential Election:Analysis of Support for Harris Factoring Pollster Differences"
+title: "A Polls-of-Polls Forecast for the 2024 United States Presidential Election Through Generalized Linear Model"
+subtitle: "The Impact of Pollster Differences, Population Type, and Poll Regency on Predicting Support for Kamala Harris"
 author: 

### 2024-11-04 paper/paper.qmd
outcome variable section edited

@@
 library(knitr)
+library(kableExtra)
 library(rstanarm)
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-- include summary statistics 
-- for every variable in your dataset that is of interest, there needs to be graphs, explanations and tables
+### The Predicted Proportion of Support a Candidate Received in a Poll (%)
+
+The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.@tbl-pct_cleaned and @fig-pct_cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
+We also get the popular vote predictions for each candidate using the predict() function and add this to the dataset as the predicted_pct variable.
+We have showed the summary statistics and distribution for predicted_pct in @tbl-pct_predicted and @fig-pct_predicted, respectively.
+
+Comparing our predictions to the data, our predictions have a smaller range than the cleaned data.
+While the difference in mean is 2\'%' lower for our prediction than our cleaned data.
+The results suggest there is less variation in our predictions than the cleaned dataset.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-pct_cleaned
+#| fig-cap: Summary statistics for the proportion of support candidates received in the poll
 #| echo: false
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'pct' variable
+summary_stats <- cleaned_president_polls %>%
+  summarize(
+    mean = round(mean(pct, na.rm = TRUE), 2),
+    median = round(median(pct, na.rm = TRUE), 2),
+    min = round(min(pct, na.rm = TRUE), 2),
+    max = round(max(pct, na.rm = TRUE), 2), 
+    sd = round(sd(pct, na.rm = TRUE), 2), 
+    n = n()
+
+  )
+
+# Create a formatted table to display summary statistics 
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 
@@
 
-Talk more about it.
+```{r}
+#| label: fig-pct_cleaned
+#| fig-cap: Distribution of the proportion of support candidates received in the poll
+#| echo: false
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Plot distribution using histogram
+ggplot(cleaned_president_polls, aes(x = pct)) + 
+  geom_histogram(binwidth = 3, fill = "green", alpha = 0.7, color = "black") +
+  theme_minimal() + 
+  labs(x = "Percentage of Support (pct)", 
+       y = "Count")
+```
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-pct_predicted
+#| fig-cap: Summary statistics for the predicted proportion of support candidates received in the poll
 #| echo: false
-#| warning: false
-#| message: false
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'pct' variable
+summary_stats <- prediction %>%
+  summarize(
+    mean = round(mean(predicted_pct, na.rm = TRUE), 2),
+    median = round(median(predicted_pct, na.rm = TRUE), 2),
+    min = round(min(predicted_pct, na.rm = TRUE), 2),
+    max = round(max(predicted_pct, na.rm = TRUE), 2), 
+    sd = round(sd(predicted_pct, na.rm = TRUE), 2), 
+    n = n()
+
+  )
+
+# Create a formatted table to display summary statistics 
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 
@@
 
-Talk way more about it.
+```{r}
+#| label: fig-pct_predicted
+#| fig-cap: Distribution of the predicted proportion of support candidates received in the poll
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
+# Plot distribution using histogram
+ggplot(prediction, aes(x = predicted_pct)) + 
+  geom_histogram(binwidth = 3, fill = "green", alpha = 0.7, color = "black") +
+  theme_minimal() + 
+  labs(x = "Predicted Percentage of Support (predicted_pct)", 
+       y = "Count")
+```
 
@@
 
+Pollster
+
+Population Type
+
+Poll Regency (old if data collected in 30 days, old to capture regent old
+
+Separated =between
+
 # Model {#sec-model}
+
 For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris.
@@
 
-The GLM is specified as follows:
-\begin{equation}
+The GLM is specified as follows: \begin{equation}
 Y_i = \beta_{0} + \beta_{1}x_{pollster_i} + \beta_{2}x_{population_i} + \beta_{3}x_{recentpoll_i} + \epsilon_{i}
@@
 # Results {#sec-results}
+
 To assess model reliability, we examined several key diagnostics.
@@
 ## Weaknesses and Next Steps {#sec-weaknesses}
+
 -   if filter thorugh state, data not collected easily
@@
 # Appendix {#sec-appendix}
+
 ## YouGov Methodology Analysis {#sec-appendix-pollster}
@@
 
-[NTD: ADD MORE DETAILS ABOUT METHODOLOGY]
-To allocate the \$100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
+\[NTD: ADD MORE DETAILS ABOUT METHODOLOGY\] To allocate the \$100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
 I would put \$50,000 into advertising and recruitment for targeted digital ads and phone surveys.
@@
 
-[NTD: ADD THIS AS ANOTHER SECTION]
-Please find the proposed question list for the online survey attached here (INSERT URL).
+\[NTD: ADD THIS AS ANOTHER SECTION\] Please find the proposed question list for the online survey attached here (INSERT URL).
 

### 2024-11-04 paper/paper.qmd
debug the compile error - fixed the lables

@@
 
-### The Predicted Proportion of Support a Candidate Received in a Poll (%)
+### The Predicted Proportion of Support a Candidate Received in a Poll
 
-The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.@tbl-pct_cleaned and @fig-pct_cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
+The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.@tbl-pct-cleaned and @fig-pct-cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
 We also get the popular vote predictions for each candidate using the predict() function and add this to the dataset as the predicted_pct variable.
-We have showed the summary statistics and distribution for predicted_pct in @tbl-pct_predicted and @fig-pct_predicted, respectively.
+We have showed the summary statistics and distribution for predicted_pct in @tbl-pct-predicted and @fig-pct-predicted, respectively.
 
 Comparing our predictions to the data, our predictions have a smaller range than the cleaned data.
-While the difference in mean is 2\'%' lower for our prediction than our cleaned data.
+While the difference in mean is 2% lower for our prediction than our cleaned data.
 The results suggest there is less variation in our predictions than the cleaned dataset.
@@
 ```{r}
-#| label: tbl-pct_cleaned
-#| fig-cap: Summary statistics for the proportion of support candidates received in the poll
+#| label: tbl-pct-cleaned
+#| tbl-cap: Summary statistics for the proportion of support candidates received in the poll
 #| echo: false
+#| message: false
+#| warning: false
 #| tbl-align: center
@@
 
-# Calculate summary statistics for the 'pct' variable
 summary_stats <- cleaned_president_polls %>%
@@
     n = n()
-
   )
 
-# Create a formatted table to display summary statistics 
 summary_stats %>%
@@
 ```{r}
-#| label: fig-pct_cleaned
+#| label: fig-pct-cleaned
 #| fig-cap: Distribution of the proportion of support candidates received in the poll
 #| echo: false
+#| message: false
+#| warning: false
 
@@
 ```{r}
-#| label: tbl-pct_predicted
-#| fig-cap: Summary statistics for the predicted proportion of support candidates received in the poll
+#| label: tbl-pct-predicted
+#| tbl-cap: Summary statistics for the predicted proportion of support candidates received in the poll
 #| echo: false
+#| message: false
+#| warning: false
 #| tbl-align: center
@@
 ```{r}
-#| label: fig-pct_predicted
+#| label: fig-pct-predicted
 #| fig-cap: Distribution of the predicted proportion of support candidates received in the poll
 #| echo: false
+#| message: false
+#| warning: false
 #| fig-width: 5 

### 2024-11-04 paper/paper.qmd
merged changes

@@
 
-Add graphs, tables and text.
+### Type of Pollster {sec-data-pollster}
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+The pollster variable was selected to consider the effect of changes in pollster on the observed variable.
+The pollster variable represents the polling organization that conducted the poll.
+The distribution of polling counts for different pollsters in @fig-pollster suggests that the data is dominated by three pollsters: Ipsos, YouGov and Siena/NYT.
+Further analysis is needed in their polling methodology to determine potential biases.
 
-Pollster
+```{r}
+#| label: tbl-pollster
+#| fig-cap: Number of unique high-quality polling organizations
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 82%
 
-Population Type
+# Calculate summary statistics for the 'pollster' variable
+summary_stats <- prediction %>%
+  summarize(
+    Pollster = n_distinct(pollster),
+  )
 
-Poll Regency (old if data collected in 30 days, old to capture regent old
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 
-Separated =between
+```
+
+```{r}
+#| label: fig-pollster
+#| fig-cap: Distribution of high-quality polling organizations
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
+# Plot distribution using histogram
+ggplot(prediction, aes(x = pollster)) +
+  geom_bar(fill = "blue", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Pollster",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+### Population Type {sec-data-populationtype}
+
+Population Type distinguishes among different groups surveyed into two groups surveyed: voters and adults.
+Adults is a more general poll that includes voters, likely voters and non-voters.
+@fig-populationtype shows that majority of the polls surveyed voters rather than the general adult population.
+The discrepancy between number of polls who surveyed voters and adults could introduce potential biases in responses, as those who are voters may have stronger opinions than those who are not.
+
+```{r}
+#| label: tbl-populationtype
+#| fig-cap: Number of unique high-quality polling organizations
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'population' variable
+summary_stats <- prediction %>%
+  summarize(
+    Population = n_distinct(population),
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-populationtype
+#| fig-cap: Distribution of population type 
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
+# Plot distribution of Population Type using histogram
+ggplot(prediction, aes(x = population)) +
+  geom_bar(fill = "orange", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Population",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+### Poll Recency {sec-data-pollrecency}
+
+Poll Recency, shown through the variable `recent_poll`, suggests that
+
+- lot more older datapoints
+
+Population Type distinguishes among different groups surveyed into two groups surveyed: voters and adults.
+Adults is a more general poll that includes voters, likely voters and non-voters.
+@fig-populationtype shows that majority of the polls surveyed voters rather than the general adult population.
+The discrepancy between number of polls who surveyed voters and adults could introduce potential baises in responses, as those who are voters may have stronger opinions than those who are not.
+
+```{r}
+#| label: tbl-recency
+#| fig-cap: Number of polls that were collected within 30 days and after
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'population' variable
+summary_stats <- prediction %>%
+  summarize(
+    Population = n_distinct(recent_poll),
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-recency
+#| fig-cap: Distribution of on the amount of polls collected in 30 days 
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
+# Plot distribution of Population Type using histogram
+ggplot(prediction, aes(x = recent_poll)) +
+  geom_bar(fill = "purple", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Poll Recency",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
 
@@
 
-\[NTD: ADD MORE DETAILS ABOUT METHODOLOGY\] To allocate the \$100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
+To allocate the $100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
 I would put \$50,000 into advertising and recruitment for targeted digital ads and phone surveys.
@@
 However, I believe \$100,000 is on the lower end to ensure a high-quality poll and will most likely cost more.
-Regardless, most of the costs will come from collecting the data, and I have budgeted for this reality (ADD CITATION).
+Regardless, most of the costs will come from collecting the data, and I have budgeted for this reality.
 
-\[NTD: ADD THIS AS ANOTHER SECTION\] Please find the proposed question list for the online survey attached here (INSERT URL).
+Please find the proposed question list for the online survey attached here (INSERT URL).
 

### 2024-11-04 paper/paper.qmd
updated predictor variable section

@@
   summarize(
-    Pollster = n_distinct(pollster),
+    count(pollster, name = "Count")
   )
@@
 summary_stats <- prediction %>%
-  summarize(
-    Population = n_distinct(population),
-  )
+  count(population, name = "Count")
 
@@
 
-Poll Recency, shown through the variable `recent_poll`, suggests that
-
-- lot more older datapoints
-
-Population Type distinguishes among different groups surveyed into two groups surveyed: voters and adults.
-Adults is a more general poll that includes voters, likely voters and non-voters.
-@fig-populationtype shows that majority of the polls surveyed voters rather than the general adult population.
-The discrepancy between number of polls who surveyed voters and adults could introduce potential baises in responses, as those who are voters may have stronger opinions than those who are not.
+Poll Recency, shown through the variable `recent_poll`, labels any polls that have been conducted in the last 30 days as Recent and any polls collected before that are labeled Older.
+Based on our results shown in @tbl-recency, there are 32 less polls collected in the last 30 days than before.
+This introduces biases in responses as older responses are more represented in our model.
+In rapidly changing political environments, public opinion can shift dramatically due to major events, debates or crises, making recent data crucial for accurate modeling.The distribution of this variable can also be found in @fig-recency.
 
@@
 
-# Calculate summary statistics for the 'population' variable
+# Calculate summary statistics for the 'recent_poll' variable
 summary_stats <- prediction %>%
-  summarize(
-    Population = n_distinct(recent_poll),
-  )
+  count(recent_poll, name = "Count")
 

### 2024-11-04 scripts/04-model_data.R
raw dataset and model code update

@@ -12,21 +12,22 @@ library(rstanarm)
 
 #### Read data ####
 # Load the cleaned data for modeling
-national_data <- read_csv("data/02-analysis_data/national_polling.csv")
+cleaned_president_polls <- read_csv("data/02-analysis_data/cleaned_president_polls.csv")
+state_data <- read_csv("data/02-analysis_data/state_polling_data.csv")
 
 #### Data filtering ####
 # Filter data to Harris estimates based on high-quality polls after she declared
-filtered_national_data <- national_data |>
+filtered_data <- cleaned_president_polls |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7
+    numeric_grade >= 3.0
   ) |>
   filter(end_date >= as.Date("2024-07-21")) # When Harris declared
 
-#### Model 1: Popular Vote Prediction ####
-popular_vote_model <- stan_glm(
+#### Model: Vote Prediction ####
+vote_for_Harris <- stan_glm(
   formula = pct ~ pollster + population + recent_poll,
-  data = filtered_national_data,
+  data = filtered_data,
   family = gaussian(),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
   prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -37,16 +38,16 @@ popular_vote_model <- stan_glm(
 )
 
 # Summarize and save the popular vote model
-summary(popular_vote_model)
-write_rds(popular_vote_model, "models/vote_for_Harris_model.rds")
-
+summary(vote_for_Harris)
+write_rds(vote_for_Harris, "models/vote_for_Harris_model.rds")
 
 #### Predictions: Popular Vote ####
-# Predicting the popular vote percentages for each candidate
-popular_vote_prediction <- predict(popular_vote_model, newdata = filtered_national_data)
-filtered_national_data <- filtered_national_data |>
+# Predicting the popular vote percentages
+popular_vote_prediction <- predict(vote_for_Harris, newdata = filtered_data)
+filtered_data <- filtered_data |>
   mutate(predicted_pct = popular_vote_prediction)
 
-# Save popular vote predictions
-write_csv(filtered_national_data, "data/02-analysis_data/popular_vote_predictions.csv")
+
+# Save predictions
+write_csv(filtered_data, "data/02-analysis_data/vote_for_Harris_predictions.csv")
 

### 2024-11-04 paper/paper.qmd
code draft for discussion

--- 2024-11-04 paper/paper.qmd
@@
 harris_model <- readRDS(file = here("models/vote_for_Harris_model.rds"))
-prediction <- read_csv(file = here("data/02-analysis_data/popular_vote_predictions.csv"), show_col_types = FALSE)
+prediction <- read_csv(file = here("data/02-analysis_data/vote_for_Harris_predictions.csv"), show_col_types = FALSE)
 ```
@@
 This test data, which includes polls not used in training, provides an unbiased estimate of model performance.
-See more details of model diagnostic here @sec-model-details-diagnostics.
+See more details of model diagnostic here: @sec-model-details-diagnostics.
 
@@
 ## How has the Forecast Changed Over Time? {#sec-first-point}
+```{r}
+filtered_prediction <- prediction %>%
+  filter(national_poll == 1)
+
+# Plot the graph with both predicted and real percentages
+ggplot(filtered_prediction, aes(x = end_date)) +
+  geom_line(aes(y = predicted_pct, color = "Predicted Percentage")) +
+  geom_line(aes(y = pct, color = "Real Percentage")) +
+  labs(
+       x = "Date",
+       y = "Percentage",
+       color = "Legend") +
+  theme_minimal()
+```
 
@@
 
-## Who is Favoured to Win Each State? {#sec-second-point}
+## Which States Prefer Harris? {#sec-second-point}
+```{r}
+state_data <- prediction %>% filter(national_poll == 0)
+# Identify missing data
+missing_data_summary <- prediction %>%
+  group_by(state) %>%
+  summarize(missing_dates = sum(is.na(predicted_pct)),
+            total_dates = n())
 
-## What are the Closest Races Across States {#sec-third-point}
+print(missing_data_summary)
+
+# Plot with points for sparse data
+ggplot(prediction, aes(x = end_date, y = predicted_pct, color = state)) +
+  geom_line(na.rm = TRUE) +
+  geom_point(na.rm = TRUE) +  # Add points to show actual data entries
+  facet_wrap(~ state, scales = "free_y") +
+  labs(
+       x = "Date",
+       y = "Predicted Percentage (%)") +
+  theme_minimal() +
+  theme(legend.position = "none",
+        strip.text = element_text(size = 10))  # Increase facet label size
+
+```
+
+```{r}
+# Filter to keep only the most recent entry for each state based on 'end_date'
+filtered_data <- prediction %>%
+  filter(!is.na(state)) %>%
+  arrange(state, end_date) %>%
+  group_by(state) %>%
+  slice_tail(n = 1) %>%
+  ungroup() %>%
+  select(state, end_date, predicted_pct)
+
+# Plot the bar chart of predicted percentage for each state with pastel colors
+ggplot(filtered_data, aes(x = reorder(state, predicted_pct), y = predicted_pct, fill = state)) +
+  geom_bar(stat = "identity") +
+  geom_text(aes(label = round(predicted_pct, 1)), hjust = -0.1) + # Display values above bars
+  labs(
+    x = "Predicted %",
+    y = "State"
+  ) +
+  theme_minimal() +
+  coord_flip() + # Flip coordinates for better readability
+  scale_fill_manual(values = scales::hue_pal()(nrow(filtered_data))) + # Pastel-like palette
+  theme(legend.position = "none") # Remove legend for simplicity
+```
 

### 2024-11-04 paper/paper.qmd
edited idealized methodology and added survey

@@
 
-Because the US operates on the Electoral College system, where state-affiliation matters, I would use Stratified Random Sampling to ensure each state is represented proportionally in the survey, improving accuracy and representativeness.
-My target population would be voter-eligible citizens in the United States who are above 18 years old.
-My sampling frame will use voter registration databases that are stratified by key demographics (age, race, gender, geographic region, income, and education level).
-Finally for my potential sample size, I will aim for around 10,000 respondents to ensure a reasonable margin of error, but I expect the actual sample to be less.
+Because the U.S.
+operates on the Electoral College system, where state-affiliation matters, I would use Stratified Random Sampling to ensure each state is represented proportionally in the survey, improving accuracy and representativeness.
+My target population would be voter-eligible citizens in the United States who are above 18 years old, with a sampling frame utilizing voter registration databases stratified by key demographics, including age, race, gender, geographic region, income, and education level.
+I aim for a sample size of around 10,000 respondents to achieve a reasonable margin of error, ensuring robust findings, though I expect the actual sample may be less due to various factors.
 
-I would primarily recruit respondents using online methods such as surveys on social media platforms like Facebook and Instagram to reach younger audiences and use interactive voice response systems on the phone to reach older or less tech-savvy demographics.
+To recruit respondents, I would primarily utilize online methods, such as surveys on social media platforms like Facebook and Instagram to engage younger audiences, and interactive voice response systems for older or less tech-savvy individuals.
+To ensure the data collected is valid, I will cross-check responses with voter registration databases and implement logic checks within the survey to detect inconsistencies.
+For instance, if a respondent claims to have already voted but indicates they’re unlikely to vote, this response will be flagged as unreliable.
 
-To ensure the data aggregated are valid, I would make sure to cross-check responses with voter registration databases and implement various logic checks within the survey to detect inconsistent answers.
-For example, if a survey respondent claims to have already voted but stated they’re unlikely to vote makes the data response unreliable.
+Addressing non-response bias is crucial, and I plan to over-sample underrepresented groups while employing multiple attempts to contact individuals online and conducting weekly polls.
+Frequent surveys will allow me to aggregate results using a moving average to smooth out short-term fluctuations in responses.
+I will also implement checks to prevent duplicate entries in the online survey and track phone responses to ensure unique participants.
+Given that the U.S.
+does not conduct elections based on a popular vote, I will weight the survey results to reflect the voting population as indicated by U.S.
+Census data and voter turnout estimates, considering demographic factors such as race, gender, and age.
 
-The biggest concern with online surveys is to handle non-response bias.
-To mitigate this, I will make sure to over-sample underrepresented groups and use multiple attempts to contact individuals online or run weekly polls.
-By running frequent surveys, I will make sure to aggregate results using a moving average in the summary results dashboard to smooth out short-term fluctuations.
-I will also make sure to develop checks online to prevent someone from filling out the form more than once or track the different people we called to ensure we do not have duplicate entries.
-Finally, because the United States does not run elections using a popular vote, I will weight the survey results to reflect the US voting population based on US Census data and voter turnout estimates, while also factoring demographic factors such as race, gender, and age.
+To allocate the \$100,000 budget, I will spend \$20,000 on a survey platform subscription, \$50,000 on targeted digital ads and phone surveys for recruitment, \$20,000 on data analysts and poll aggregation services, and \$10,000 for re-contact surveys and miscellaneous costs.
+However, I recognize that this budget may be on the lower end to ensure a high-quality poll and that the actual costs might exceed this amount.
+Most expenses will arise from data collection, and I have accounted for this in my budget planning.
 
-To allocate the $100,000 budget, I would budget \$20,000 for a survey platform subscription to handle a large amount of data.
-I would put \$50,000 into advertising and recruitment for targeted digital ads and phone surveys.
-I will also use \$20,000 for data analysts and poll aggregation services and \$10,000 for re-contact surveys, contingency costs, and other miscellaneous costs.
-However, I believe \$100,000 is on the lower end to ensure a high-quality poll and will most likely cost more.
-Regardless, most of the costs will come from collecting the data, and I have budgeted for this reality.
-
-Please find the proposed question list for the online survey attached here (INSERT URL).
+Please find the proposed question list for the online survey attached [here](https://forms.gle/AFqodUEk48o864aN6).
 

### 2024-11-04 paper/paper.qmd
update on model coefficient
modified predictor variables

@@
 
-To ensure high data quality, we filtered the dataset to include polls with a numeric grade of 2.7 or above and focused only on polls conducted after July 21, 2024, the date of Harris' declaration of candidacy.
+To ensure high data quality, we filtered the dataset to include polls with a numeric grade of 3.0 and focused only on polls conducted after July 21, 2024, the date of Harris' declaration of candidacy.
 The filter helped limit bias from older polls from when Joe Biden was the Democratic nominee, which may not reflect the current voter sentiment.
@@
 
-The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.@tbl-pct-cleaned and @fig-pct-cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
+The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.
+@tbl-pct-cleaned and @fig-pct-cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
 We also get the popular vote predictions for each candidate using the predict() function and add this to the dataset as the predicted_pct variable.
@@
 
-### Type of Pollster {sec-data-pollster}
+### Type of Pollster {#sec-data-pollster}
 
@@
 The pollster variable represents the polling organization that conducted the poll.
-The distribution of polling counts for different pollsters in @fig-pollster suggests that the data is dominated by three pollsters: Ipsos, YouGov and Siena/NYT.
+The distribution of polling counts for different pollsters in @fig-pollster suggests that the data is dominated by two pollsters: YouGov and Siena/NYT.
 Further analysis is needed in their polling methodology to determine potential biases.
@@
 #| label: tbl-pollster
-#| fig-cap: Number of unique high-quality polling organizations
+#| tbl-cap: Number of unique high-quality polling organizations
 #| echo: false
@@
 summary_stats <- prediction %>%
-  summarize(
   count(pollster, name = "Count")
-  )
 
@@
        y = "Count") +
+  coord_flip() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
@@
 
-### Population Type {sec-data-populationtype}
+### Population Type {#sec-data-populationtype}
 
@@
 #| label: tbl-populationtype
-#| fig-cap: Number of unique high-quality polling organizations
+#| tbl-cap: Number of unique high-quality polling organizations
 #| echo: false
@@
 
-### Poll Recency {sec-data-pollrecency}
+### Poll Recency {#sec-data-pollrecency}
 
-Poll Recency, shown through the variable `recent_poll`, labels any polls that have been conducted in the last 30 days as Recent and any polls collected before that are labeled Older.
-Based on our results shown in @tbl-recency, there are 32 less polls collected in the last 30 days than before.
-This introduces biases in responses as older responses are more represented in our model.
-In rapidly changing political environments, public opinion can shift dramatically due to major events, debates or crises, making recent data crucial for accurate modeling.The distribution of this variable can also be found in @fig-recency.
+Poll Recency, shown through the variable `recent_poll`, labels any polls conducted in the last 30 days as Recent and any polls collected before that are labeled Older.
+Based on our results shown in @tbl-recency, 19 more polls have been collected in the last 30 days than before.
+This introduces biases in responses as recent responses are more represented in our model.
+In rapidly changing political environments, public opinion can shift dramatically due to major events, debates, or crises, making recent data crucial for accurate modeling.
+The distribution of this variable can also be found in @fig-recency.
 
@@
 #| label: tbl-recency
-#| fig-cap: Number of polls that were collected within 30 days and after
+#| tbl-cap: Number of polls that were collected within 30 days and after
 #| echo: false
@@
 Here, we utilized high-quality national polling data gathered after Harris’s campaign announcement.
-We filtered the dataset to include polls with a numeric grade of 2.7 or above, ensuring data quality, and focused only on polls conducted after July 21, 2024, the date of Harris’s declaration of candidacy.
+We filtered the dataset to include polls with a numeric grade of 3.0, ensuring data quality, and focused only on polls conducted after July 21, 2024, the date of Harris’s declaration of candidacy.
 This filter helps limit bias from older polls that may not reflect the current voter sentiment.
@@
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
-For example, the 'CES / YouGov' pollster shows a positive impact, while 'Siena/NYT' has a slight negative effect.
+For example, the 'populationVoters' shows a positive impact, while most of the pollsters has a slight negative effect.
 These error bars help contextualize the reliability of each predictor.

### 2024-11-04 paper/paper.qmd
updated formatting of graphs in discussion

@@
 library(knitr)
-library(kableExtra)
+library(kable)
 library(rstanarm)
@@
 # Display the summary statistics in a nicely formatted table
-summary_stats %>%
+summary_stats |>
+  kable
   kable() %>%
@@
 ## How has the Forecast Changed Over Time? {#sec-first-point}
+
 ```{r}
+#| label: fig-predicted-and-real
+#| fig-cap: Comparison of Predicted vs Actual National Poll Percentages for Harris Over Time
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
 filtered_prediction <- prediction %>%
@@
 
-asdfasdfasdfasdf
+## Which States Prefer Harris? {#sec-second-point}
 
-## Which States Prefer Harris? {#sec-second-point}
 ```{r}
+#| label: tbl-state-missing
+#| fig-cap: Summary of missing and total dates across key states
+#| echo: false
+#| tbl-align: center
+
 state_data <- prediction %>% filter(national_poll == 0)
@@
 
-print(missing_data_summary)
+# Display the summary statistics in a nicely formatted table
+missing_data_summary %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-state-comparison
+#| fig-cap: State-by-State comparison of predicted percentages over time
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
 
@@
 ```{r}
+#| label: fig-state-summary
+#| fig-cap: Summary of Kamala Harris' predicted popular vote by key states
+#| echo: false
+#| fig-width: 5 
+#| fig-align: center
+
 # Filter to keep only the most recent entry for each state based on 'end_date'
@@
 ## Implications {#sec-implications}
+
+asdfasdfasdf
 

### 2024-11-04 paper/paper.qmd
fix error: Error in parse()

@@
 library(knitr)
-library(kable)
 library(rstanarm)
@@
 # Display the summary statistics in a nicely formatted table
-summary_stats |>
-  kable
+summary_stats %>%
   kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
 

### 2024-11-04 paper/paper.qmd
added discussion, abstract, and a citation

@@
 title: "A Polls-of-Polls Forecast for the 2024 United States Presidential Election Through Generalized Linear Model"
-subtitle: "The Impact of Pollster Differences, Population Type, and Poll Regency on Predicting Support for Kamala Harris"
+subtitle: "The Impact of Pollster Differences, Population Type, and Poll Recency on Predicting Support for Kamala Harris"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The 2024 U.S.
+Presidential Election has far-reaching global implications for economic, environmental, and social policy, making it a focal point for international attention.
+This paper applies a Bayesian generalized linear model to analyze trends in Kamala Harris' support, accounting for pollster differences, population types, and poll recency.
+Despite steady support at 48%, Harris is projected to lose the national popular vote and key battleground states to Donald Trump.
+Our findings emphasize the significance of polling methodologies and their role in predicting U.S.
+policy outcomes.
+Future research should incorporate the Electoral College to address current limitations."
 format: pdf
@@
 Factoring in the different results of poll-making organizations, our model found that specific pollsters or demographic characteristics increase or decrease Harris' predicted voter share.
+Analyzing the data at a national and state level, we predict that Harris has a minority in the national popular vote and in key swing states, leading us to conclude that Donald Trump will win the 2024 US Presidential Election.
 Understanding the trajectory of the US might make in foreign policy, economic or global politics prediction allows global stakeholders to take preemptive measures for policy changes of the newly elected government.
@@
 
+\newpage
+
 ```{r}
@@
 
+\newpage
+
 ### Population Type {#sec-data-populationtype}
@@
 ```
-
+\newpage
 ```{r}
@@
 ```
+\newpage
 
@@
 ```
+
+\newpage
 
@@
 
+@fig-predicted-and-real maps the predicted and real approval percentages for Kamala Harris during the time of her campagin.
+In @fig-predicted-and-real, our model shows that predicted approval ratings during Kamala Harris' campagin sat at around 47%, on average, which is in line with current ABC polling methods.
+However, if we examine the real polling approval rating percentages, it shows that Kamala Harris' approval rating across polls significantly varied over time, but remained below 50%.
+The observed proportion of support over time varies because different polls have different methodlogies and suggests that there was significant variation in Kamala Harris' polling results during her campaign.
+
+At the beginning of her campagin, Harris' approval rating significantly rose, but it eventually leveled out to just around 47%.
+Her initial rise in the polls came from excitement from the replacement of Joe Biden and the added momentum from a new candidate.
+Evidently, leading up to the election, Harris' approval ratings saw spikes and dips, most likely due to key catalysts such as the Presidential Debate, Hurricane Milton, and the escalation in the Ukraine-Russian War and the Israel-Palestine Conflict.
+However, leading up to the election, @fig-predicted-and-real shows Harris' approval rating settling just approximately at 47%, signalling that Harris is likely to lose the popular vote, or at very best, the 2024 US presidential election will be extremeley close.
+
+The most apparent observation from @fig-predicted-and-real is that Harris' approval rating is quite volatile over time.
+The reason for this volatility is due to the differences between pollsters' methodology, leading to drastically different results.
+Our model factors for these differences and notes the different pollsters as a notable predictor variable.
+The apprarent variation in @fig-predicted-and-real suggests the influence methodology and time have on the results of the poll.
+As time progressed, Harris' approval rating also varied, signalling that any candidate's momentum can change rapidly and maintaining a stable base of support can be challenging in a dynamic environment.
+
+Overall, @fig-predicted-and-real evaluates Harris' approval rate throughout the duration of her campagin.
+Evaluating the real results and predicted percentages from the pollsters and models, respectively, Harris' is at a disadvantage heading into the election, with Donald Trump expected to win the popular vote.
+
 ```{r}
@@
 ## Which States Prefer Harris? {#sec-second-point}
+
+However, unlike other countries, the United States' election runs through the Electoral College.
+To increase representation for smaller states, the Electoral College led to the formation of "swing states", which are states that have similar levels of support among voters and important to the outcome of the election.
+Notable swing states include Wisconsin, Pennsylvania, Arizona, Minnnesota, Georgia, and Florida.
+Therefore, to have a clearer picture on the outcome of the United States election, it is important to do a state-by-state analysis to understand Harris' approval ratings across key states.
+
+In our model, by separating for state-specific polls, we are able to analyze and model out the poll results by states where state-specific polls were conducted.
+Examining @fig-state-comparison, common battle ground states such as Arizona, Pennsylvania, Wisconsin and Michigan are polled more frequently at the state-level than partisan states such as California and New York.
+Examining the results at these key battleground states, the results show that Harris' popularity recenty fell below 50% during October and November in key states such as Arizona, Michigan, Pennsylvania, Nevada and Wisconsin.
+Though Harris has predicted to have a majority in Ohio and Florida, the disappointing results in the other states make sher lead in these states less signifiant.
+Moreover, looking at the general NA category,representing all non-state specific polls, Harris' predicted approval rating hovers at 48%, which means she is expected to lose the popular vote to Trump.
+
+@fig-state-summary shows that Harris is projected to have a minor lead with 50.4% of the popular vote in five key states: Virginia, Ohio, New Hampshire, Minnesota, and Florida.
+However, in the other states, she is expected to lose the popular vote.
+States such as Pennsylvania, Nevada, Michigan, and Georgia are critical states in order to win the election.
+As Harris has a noticeable projected minority in these states, she not only is expected to lose the popular vote, but Trump has a projected lead over Harris in these battleground states, hurting her chances at being elected.
+
+Overall, our model predicts that Harris is unlikely to win the general popular vote or key swing states. Consequently, we anticipate Donald Trump will win the 2024 U.S. Presidential election.
+
+\newpage
 
@@
 #| echo: false
-#| fig-width: 5 
 #| fig-align: center
@@
   theme(legend.position = "none",
-        strip.text = element_text(size = 10))  # Increase facet label size
+        strip.text = element_text(size = 8))  # Increase facet label size
 
@@
 
-asdfasdfasdf
+If our prediction is correct and Trump returns to office, we would see a shift in the United States' domestic and foreign policy compared to the Biden administration.
+Donald Trump's past presidency from 2016 to 2020 was marked with tax cuts, stronger border enforcement, isolationist foreign policy, and a massive rollback on environmental regulations [@BBCNews2024].
+During this campaign cycle, Trump has largely promoted similar policies, representing a stark contrast to Harris' campaign.
+The implication of the paper is that if Trump wins the election as projected, then we would see the re-introduction of these policies and a different America on the global stage.
+As the United States election is one of the most important international events of this year, predicting the event's outcome is critical for foreign governments and businesses to guide strategy on US relations and regulations.
+
+Moreover, the paper has significant lessons for the Harris campaign.
+Harris should prioritize swing states where small shifts in public opinion can determine electoral outcomes.
+Harris' campagin may need to adjust its messaging and outreach in battleground states to mitigate her projected losses.
+Also, shown through @fig-predicted-and-real, popularity can flucturate based on public enthusiasm and external factors.
+Campagins must focus on mobilizing their voter base effectively in the weeks leading up to the election.
+Though the paper can help guide the Harris campaign leading up to the election, it also can serve to help future candidates in guiding their strategy surrounding shaping public opinion and prioritizing swing states.
 
@@
 
--   if filter thorugh state, data not collected easily
+The first weakness in our study is the discrepancy of datapoints between states.
+@fig-state-summary shows that certain states have more frequent state-specific polls than others.
+While states like California, Montana, and New York have historically leaned towards one party than the other, the discrepancy in datapoints between states is a weakness in our data and hurts our abilty to draw conclusions from our data, as it is difficult to make out trends.
+Therefore, for future studies, a next potential step could be to conduct more state-specific polls and map out a candidate's popularity across each state.
 
-Weaknesses and next steps should also be included.
+The second weakness of the study is its reliance on the popular vote, which may not accurately reflect the results of the election.
+The United States does not elect its President based on the popular vote, but it uses the Electoral College.
+While the popular vote is a valuable metric, the Electoral College system means that a candidate can lose the popular vote and still win the presidency, as seen in past elections.
+Harris' projected loss in key states, combine dwith a predicted minority in the national polls, signals that simply gaining a majority in the popular vote may not be sufficient for electoral success under the current system.
+Therefore, the next step for this study is to further incorporate the Electoral College and its methodology to the study, requiring more state-specific polling and adjusting the estimand.
+
+Finally, as we are working with polling data, there is potential for bias.
+As mentioned earlier in @sec-data-pollrecency, potential for bias may disproportionately influence model predictions.
+This has broader implications for the accuracy of electoral forecasts and highlights the need for continually updated polling to provide a more accurate reflection of voter sentiment as Election Day approaches.
 

### 2024-11-04 paper/paper.qmd
minor grammar edits

@@
 Examining the results at these key battleground states, the results show that Harris' popularity recenty fell below 50% during October and November in key states such as Arizona, Michigan, Pennsylvania, Nevada and Wisconsin.
-Though Harris has predicted to have a majority in Ohio and Florida, the disappointing results in the other states make sher lead in these states less signifiant.
+Though Harris has predicted to have a majority in Ohio and Florida, the disappointing results in the other states makes her lead in these states less signifiant.
 Moreover, looking at the general NA category,representing all non-state specific polls, Harris' predicted approval rating hovers at 48%, which means she is expected to lose the popular vote to Trump.
@@
 However, in the other states, she is expected to lose the popular vote.
-States such as Pennsylvania, Nevada, Michigan, and Georgia are critical states in order to win the election.
+Pennsylvania, Nevada, Michigan, and Georgia are critical states for determining the outcome of the election.
 As Harris has a noticeable projected minority in these states, she not only is expected to lose the popular vote, but Trump has a projected lead over Harris in these battleground states, hurting her chances at being elected.
 
-Overall, our model predicts that Harris is unlikely to win the general popular vote or key swing states. Consequently, we anticipate Donald Trump will win the 2024 U.S. Presidential election.
+Thus, our model predicts that Harris is unlikely to win the general popular vote or key swing states. Consequently, we anticipate Donald Trump will win the 2024 U.S. Presidential election.
 

### 2024-11-04 paper/paper.qmd
update on unused dataset

@@
 cleaned_president_polls <-read_csv(file = here("data/02-analysis_data/cleaned_president_polls.csv"), show_col_types = FALSE)
-cleaned_data_survey <- read_csv(file = here("data/02-analysis_data/national_polling.csv"), show_col_types = FALSE)
-state_polling_data <- read_csv(file = here("data/02-analysis_data/state_polling_data.csv"), show_col_types = FALSE)
-state_analysis <- read_csv(file = here("data/02-analysis_data/state_analysis.csv"), show_col_types = FALSE)
 

### 2024-11-04 scripts/02-clean_data.R

@@ -70,38 +70,6 @@ cleaned_data <-
   # Remove all rows with any NA values except for the state (which is already handled)
   drop_na(pct, sample_size, candidate_name, pollster)
 
-#### Create datasets for national and state polls ####
-# National polling data
-national_polling_data <- 
-  cleaned_data |>
-  filter(national_poll == 1) |>
-  select(-state, -national_poll)
-
-# State-level polling data
-state_polling_data <- 
-  cleaned_data |>
-  filter(national_poll == 0) |>
-  select(-national_poll)
-
-
-# State-level analysis for electoral votes and close races
-state_analysis <- 
-  state_polling_data |>
-  group_by(state, candidate_name, end_date) |>
-  ungroup() |>
-  group_by(state) |>
-  # Determine who's leading in each state
-  mutate(
-    winner = candidate_name[which.max(pct)],
-    closest_race = abs(diff(range(pct)))  # The difference between the top two candidates' avg_pct
-  ) |>
-  arrange(closest_race) |>  # Sort by closest races
-  select(-candidate_name, -numeric_grade, -sample_size, -population, -start_date, -recent_poll, -pollster, - party)
-
-
 # Export cleaned datasets
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_president_polls.csv")
-write_csv(national_polling_data, "data/02-analysis_data/national_polling.csv")
-write_csv(state_polling_data, "data/02-analysis_data/state_polling_data.csv")
-write_csv(state_analysis, "data/02-analysis_data/state_analysis.csv")
 

### 2024-11-04 paper/paper.qmd
added figure height toggle

@@
 Further analysis is needed in their polling methodology to determine potential biases.
-
-\newpage
 
@@
 #| fig-width: 5
+#| fig-height: 5
 #| fig-align: center
@@
 #| fig-width: 5 
+#| fig-height: 5
 #| fig-align: center
@@
 #| label: tbl-state-missing
-#| fig-cap: Summary of missing and total dates across key states
+#| tbl-cap: Summary of missing and total dates across key states
 #| echo: false
@@
 #| echo: false
+#| message: false
 #| fig-align: center

### 2024-11-04 paper/paper.qmd
update on paper - mostly check grammar, wording and writing

--- 2024-11-04 paper/paper.qmd
@@
 The United States Presidential Election is one of the most consequential events of 2024.
-As a key country in international relations and the largest economy in the world, the results of the U.S election determines the country's domestic and foreign policy for the next four years, which will have a significant impact on important initiatives such as tackling climate change or international conflicts.
-In anticipation of the 2024 U.S election, this paper aims to predict the possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.
+As a key country in international relations and the largest economy in the world, the results of the U.S.
+election determine the country's domestic and foreign policy for the next four years, which will have a significant impact on important initiatives such as tackling climate change or international conflicts.
+In anticipation of the 2024 U.S.
+election, this paper aims to predict the possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.
 
 We forecast support for Kamala Harris based on polling results and apply a Bayesian generalized linear model.
-The main parameter of interest is the proportion of vote or support that Harris received in surveys, which is traced over time.
-By considering the results from different poll-making organizations and other demographic factors, our objective is to correct for variation across different voter bases with
+The main parameter of interest is the proportion of votes or support that Harris received in surveys, which is traced over time.
+By considering the results from different poll-making organizations and other demographic factors, our objective is to correct for variation across different voter bases.
 
 Factoring in the different results of poll-making organizations, our model found that specific pollsters or demographic characteristics increase or decrease Harris' predicted voter share.
-Analyzing the data at a national and state level, we predict that Harris has a minority in the national popular vote and in key swing states, leading us to conclude that Donald Trump will win the 2024 US Presidential Election.
-Understanding the trajectory of the US might make in foreign policy, economic or global politics prediction allows global stakeholders to take preemptive measures for policy changes of the newly elected government.
+Analyzing the data at a national and state level, we predict that Harris has a minority in the national popular vote and key swing states, leading us to conclude that Donald Trump will win the 2024 US Presidential Election.
+Understanding the trajectory the US might make in foreign policy, economic, or global politics prediction allows global stakeholders to take preemptive measures for policy changes of the newly elected government.
 Thus, this study is a strong tool to navigate uncertainty around the 2024 election.
@@
 Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.
-The dataset includes a wide range of poll results from various polls, with keyv ariables such as pollster, sample size, percentage of support for Harris, and the date conducted for polls.
+The dataset includes a wide range of poll results from various polls, with key variables such as pollster, sample size, percentage of support for Harris, and the date conducted for polls.
 
-To ensure high data quality, we filtered the dataset to include polls with a numeric grade of 3.0 and focused only on polls conducted after July 21, 2024, the date of Harris' declaration of candidacy.
-The filter helped limit bias from older polls from when Joe Biden was the Democratic nominee, which may not reflect the current voter sentiment.
-
-In performing the analysis, we used several R pacakages.
+In performing the analysis, we used several R packages.
 We used tidyverse [@tidyverse] for data manipulation and visualization, and rstanarm [@rstanarm] for Bayesian modeling.
@@
 
-FiveThirtyEight aggregates various poll results from national and state-wide polls, showing 16867 observations [@fivethirtyeight_polls].
+FiveThirtyEight aggregates various poll results from national and state-wide polls, showing 17765 observations [@fivethirtyeight_polls] as of November 3, 2024.
 The poll takes a sample of the electorate and asks for the voters' candidate of choice.
-By factoring polls from a state and national level, FiveThirtyEight aims to represnet public perception on the two candidates during the election to predict the election's outcome.
+By factoring polls from a state and national level, FiveThirtyEight aims to represent the public perception of the two candidates during the election to predict the election's outcome.
 
@@
 
-Beyond filtering for high-quality pollsters, we filtered for various criteria to prepare the analysis data.From the raw data, we filtered out any missing numbers, cleaned names, and removed all polls on non-Democrat or Republican candidates, effectively on comparing Harris and Trump.
-After determining state-level analysis for electoral votes and close races, we created different datasets for national and state level polls.
-Regardless, selection bias and sampling bias is still a concern because polls only represent a part of the population.
-Finally, polling methodology difference can introduce bias in the study.
-Overall, we use the data from individual responses to election polls, and filter the data for analysis to make a prediction on who will win the popular vote based on data from the pollsters.
+Beyond filtering for high-quality pollsters, we filtered for various criteria to prepare the analysis data.
+From the raw data, we filtered out any missing numbers, cleaned names, and removed all polls other than Harris, effectively generating prediction support for Harris.
+Our dataset contains qualified polls for both national and state-level polls.
+Regardless, selection bias and sampling bias are still a concern because polls only represent a part of the population.
+Finally, polling methodology differences can introduce bias in the study.
+Overall, we use the data from individual responses to election polls and filter the data for analysis to predict who will win the popular vote based on data from the pollsters.
 
@@
 The main variable we aim to forecast is the pct variable, which represents the proportion of the vote a candidate received in a poll.
-@tbl-pct-cleaned and @fig-pct-cleaned shows the summary statistics and distribution of pct variable in a filtered dataset that only comprises of votes from relatively high-quality polling organizations.
+@tbl-pct-cleaned and @fig-pct-cleaned show the summary statistics and distribution of the pct variable in a filtered dataset that only comprises votes from relatively high-quality polling organizations.
 We also get the popular vote predictions for each candidate using the predict() function and add this to the dataset as the predicted_pct variable.
-We have showed the summary statistics and distribution for predicted_pct in @tbl-pct-predicted and @fig-pct-predicted, respectively.
-
+We have shown the summary statistics and distribution for predicted_pct in @tbl-pct-predicted and @fig-pct-predicted, respectively.
 Comparing our predictions to the data, our predictions have a smaller range than the cleaned data.
-While the difference in mean is 2% lower for our prediction than our cleaned data.
-The results suggest there is less variation in our predictions than the cleaned dataset.
+The difference in the mean is 2% lower for our prediction than our cleaned data.
+The results suggest there is less variation in our predictions than in the cleaned dataset.
 
 ```{r}
-#| label: tbl-pct-cleaned
-#| tbl-cap: Summary statistics for the proportion of support candidates received in the poll
 #| echo: false
@@
 #| warning: false
+#| label: tbl-pct-cleaned
+#| tbl-cap: Summary statistics for the proportion of support candidates received in the poll
 #| tbl-align: center
@@
 ```{r}
-#| label: fig-pct-cleaned
-#| fig-cap: Distribution of the proportion of support candidates received in the poll
 #| echo: false
@@
 #| warning: false
+#| label: fig-pct-cleaned
+#| fig-cap: Distribution of the proportion of support candidates received in the poll
 
@@
 ```{r}
-#| label: tbl-pct-predicted
-#| tbl-cap: Summary statistics for the predicted proportion of support candidates received in the poll
 #| echo: false
@@
 #| warning: false
+#| label: tbl-pct-predicted
+#| tbl-cap: Summary statistics for the predicted proportion of support candidates received in the poll
 #| tbl-align: center
@@
 ```{r}
-#| label: fig-pct-predicted
-#| fig-cap: Distribution of the predicted proportion of support candidates received in the poll
 #| echo: false
@@
 #| warning: false
+#| label: fig-pct-predicted
+#| fig-cap: Distribution of the predicted proportion of support candidates received in the poll
 #| fig-width: 5 
@@
 
-The pollster variable was selected to consider the effect of changes in pollster on the observed variable.
+The pollster variable was selected to consider the effect of changes in the pollster on the observed variable.
 The pollster variable represents the polling organization that conducted the poll.
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: tbl-pollster
 #| tbl-cap: Number of unique high-quality polling organizations
-#| echo: false
 #| tbl-align: center
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-pollster
 #| fig-cap: Distribution of high-quality polling organizations
-#| echo: false
 #| fig-width: 5 
@@
 
-Population Type distinguishes among different groups surveyed into two groups surveyed: voters and adults.
-Adults is a more general poll that includes voters, likely voters and non-voters.
-@fig-populationtype shows that majority of the polls surveyed voters rather than the general adult population.
-The discrepancy between number of polls who surveyed voters and adults could introduce potential biases in responses, as those who are voters may have stronger opinions than those who are not.
+Population Type was distinguished among different groups surveyed into two groups surveyed: 'Voters' and 'Adults'.
+'Adults' is a more general poll that includes voters, likely voters, and non-voters.
+@fig-populationtype shows that the majority of the polls surveyed voters rather than the general adult population.
+The discrepancy between the number of polls that surveyed voters and adults could introduce potential biases in responses, as those who are voters may have stronger opinions than those who are not.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: tbl-populationtype
 #| tbl-cap: Number of unique high-quality polling organizations
-#| echo: false
 #| tbl-align: center
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-populationtype
 #| fig-cap: Distribution of population type
-#| echo: false
 #| fig-width: 5 
@@
 
-Poll Recency, shown through the variable `recent_poll`, labels any polls conducted in the last 30 days as Recent and any polls collected before that are labeled Older.
+Poll Recency, shown through the variable `recent_poll`, labels any polls conducted in the last 30 days as Recent, and any polls collected before that are labeled Older.
 Based on our results shown in @tbl-recency, 19 more polls have been collected in the last 30 days than before.
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: tbl-recency
 #| tbl-cap: Number of polls that were collected within 30 days and after
-#| echo: false
 #| tbl-align: center
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-recency
 #| fig-cap: Distribution of on the amount of polls collected in 30 days 
-#| echo: false
 #| fig-width: 5 
@@
 This approach allows us to capture polling characteristics and account for known variations between pollsters, population types, and the recency of polling data.
-By incorporating these factors, we aim to model a nuanced estimate of Harris's projected popular vote.
+By incorporating these factors, we aim to model an estimate of support for Harris.
 
@@
 We filtered the dataset to include polls with a numeric grade of 3.0, ensuring data quality, and focused only on polls conducted after July 21, 2024, the date of Harris’s declaration of candidacy.
-This filter helps limit bias from older polls that may not reflect the current voter sentiment.
+The filter helped limit bias from older polls from when Joe Biden was the Democratic nominee, which may not reflect the current voter sentiment.
 
@@
 Population Type distinguishes among different groups surveyed (e.g., voters, likely voters) and helps in capturing generalizability.
-Categorizing polls as either recent or not ensures that more recent polls, which are better predictors of voting behavior closer to the election, receive appropriate emphasis.
+Categorizing polls as either recent or older ensures that more recent polls are better predictors of voting behavior closer to the election.
 $Y_i$ denotes the predicted popular vote percentage for Kamala Harris in the $i$-th poll.
@@
 
-The selection of predictors in this model is based on observable patterns in the data, such as the consistent tendencies of certain pollsters and the impact of population type on polling outcomes.
-To enhance the model's robustness, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded in previous election data and polling analysis.
+To enhance the model, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded.
 For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating overfitting.
-Similarly, the intercept uses a normal prior with a mean of 0 and scale of 2.5 to stabilize model estimates.
+Similarly, the intercept uses a normal prior with a mean of 0 and a scale of 2.5 to stabilize model estimates.
 For the error term (sigma), an exponential prior with a rate of 1 was selected to constrain the residuals, aligning with Gaussian assumptions.
-These priors offer a balance between model flexibility and constraint, informed by established trends in polling data and comparable electoral forecasts.
 
 The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
-Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for each candidate, using filtered national data.
-These predictions are then added to the dataset as a new variable (`predicted_pct`) for further analysis.
-The results are saved in a CSV file (`popular_vote_predictions.csv`), enabling an in-depth examination of potential outcomes in the popular vote.
-This approach facilitates understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on demographic and polling data.
+Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for Harris.
+This prediction is then added to the dataset as a new variable (`predicted_pct`) for further analysis.
+The results are saved in a CSV file (`vote_for_Harris_predictions.csv`), enabling an examination of potential outcomes in the vote.
+This approach facilitates an understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on polling data.
 
@@
 To assess model reliability, we examined several key diagnostics.
-Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence.
+Convergence metrics, such as Rhat values, were very close to 1 for all parameters, indicating strong convergence.
 Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability.
-For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy.
-This test data, which includes polls not used in training, provides an unbiased estimate of model performance.
 See more details of model diagnostic here: @sec-model-details-diagnostics.
 
-The model operates under the assumption that residuals follow a Gaussian distribution, though this may not entirely capture extreme polling variances.
-Additionally, pollster effects are treated as fixed, which, while limiting the model’s ability to reflect dynamic shifts in polling methodologies, simplifies the model and reduces the risk of overfitting.
-Alternative specifications, including a hierarchical model with random intercepts for pollsters, were considered.
-However, the added complexity of this approach did not much improve predictive accuracy.
-The chosen GLM specification thus offered a balanced approach, optimizing both interpretability and performance, making it the preferred model for our study.
+The model assumes residuals follow a Gaussian distribution, though this may not fully capture extreme polling variances.
+Pollster effects are treated as fixed to simplify the model and reduce overfitting, even if it limits responsiveness to shifts in polling methods.
+A hierarchical model with random pollster intercepts was considered, but its added complexity did not enhance accuracy.
+The selected GLM specification balances interpretability and performance, making it the preferred choice for this study.
 
@@
 Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share.
-For example, the 'populationVoters' shows a positive impact, while most of the pollsters has a slight negative effect.
+For example, the 'populationVoters' shows a positive impact, while most of the pollsters have a slight negative effect.
 These error bars help contextualize the reliability of each predictor.
@@
 
-@fig-predicted-and-real maps the predicted and real approval percentages for Kamala Harris during the time of her campagin.
-In @fig-predicted-and-real, our model shows that predicted approval ratings during Kamala Harris' campagin sat at around 47%, on average, which is in line with current ABC polling methods.
+@fig-predicted-and-real maps the predicted and real approval percentages for Kamala Harris during the time of her campaign.
+In @fig-predicted-and-real, our model shows that predicted approval ratings during Kamala Harris' campaign sat at around 47%, on average, which is in line with current ABC polling methods.
 However, if we examine the real polling approval rating percentages, it shows that Kamala Harris' approval rating across polls significantly varied over time, but remained below 50%.
-The observed proportion of support over time varies because different polls have different methodlogies and suggests that there was significant variation in Kamala Harris' polling results during her campaign.
+The observed proportion of support over time varies because different polls have different methodologies and suggest that there was significant variation in Kamala Harris' polling results during her campaign.
 
-At the beginning of her campagin, Harris' approval rating significantly rose, but it eventually leveled out to just around 47%.
+At the beginning of her campaign, Harris' approval rating significantly rose, but it eventually leveled out to just around 47%.
 Her initial rise in the polls came from excitement from the replacement of Joe Biden and the added momentum from a new candidate.
-Evidently, leading up to the election, Harris' approval ratings saw spikes and dips, most likely due to key catalysts such as the Presidential Debate, Hurricane Milton, and the escalation in the Ukraine-Russian War and the Israel-Palestine Conflict.
-However, leading up to the election, @fig-predicted-and-real shows Harris' approval rating settling just approximately at 47%, signalling that Harris is likely to lose the popular vote, or at very best, the 2024 US presidential election will be extremeley close.
+Leading up to the election, Harris' approval ratings saw spikes and dips, most likely due to key catalysts such as the Presidential Debate, Hurricane Milton, and the escalation in the Ukraine-Russian War and the Israel-Palestine Conflict.
+However, leading up to the election, @fig-predicted-and-real shows Harris' approval rating settling just approximately at 47%, signaling that Harris is likely to lose the popular vote, or at very best, the 2024 US presidential election will be close.
 
 The most apparent observation from @fig-predicted-and-real is that Harris' approval rating is quite volatile over time.
-The reason for this volatility is due to the differences between pollsters' methodology, leading to drastically different results.
+The reason for this volatility is due to the differences between pollsters' methodologies, leading to drastically different results.
 Our model factors for these differences and notes the different pollsters as a notable predictor variable.
-The apprarent variation in @fig-predicted-and-real suggests the influence methodology and time have on the results of the poll.
-As time progressed, Harris' approval rating also varied, signalling that any candidate's momentum can change rapidly and maintaining a stable base of support can be challenging in a dynamic environment.
-
-Overall, @fig-predicted-and-real evaluates Harris' approval rate throughout the duration of her campagin.
-Evaluating the real results and predicted percentages from the pollsters and models, respectively, Harris' is at a disadvantage heading into the election, with Donald Trump expected to win the popular vote.
+The apparent variation in @fig-predicted-and-real suggests the influence methodology and time have on the results of the poll.
+As time progressed, Harris' approval rating also varied, signaling that any candidate's momentum can change rapidly and that maintaining a stable base of support can be challenging in a dynamic environment.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-predicted-and-real
 #| fig-cap: Comparison of Predicted vs Actual National Poll Percentages for Harris Over Time
-#| echo: false
 #| fig-width: 5 
@@
 However, unlike other countries, the United States' election runs through the Electoral College.
-To increase representation for smaller states, the Electoral College led to the formation of "swing states", which are states that have similar levels of support among voters and important to the outcome of the election.
-Notable swing states include Wisconsin, Pennsylvania, Arizona, Minnnesota, Georgia, and Florida.
-Therefore, to have a clearer picture on the outcome of the United States election, it is important to do a state-by-state analysis to understand Harris' approval ratings across key states.
+To increase representation for smaller states, the Electoral College led to the formation of "swing states", which are states that have similar levels of support among voters and are important to the outcome of the election.
+Notable swing states include Wisconsin, Pennsylvania, Arizona, Minnesota, Georgia, and Florida.
+Therefore, to have a clearer picture of the outcome of the United States election, it is important to do a state-by-state analysis to understand Harris' approval ratings across key states.
 
-In our model, by separating for state-specific polls, we are able to analyze and model out the poll results by states where state-specific polls were conducted.
-Examining @fig-state-comparison, common battle ground states such as Arizona, Pennsylvania, Wisconsin and Michigan are polled more frequently at the state-level than partisan states such as California and New York.
-Examining the results at these key battleground states, the results show that Harris' popularity recenty fell below 50% during October and November in key states such as Arizona, Michigan, Pennsylvania, Nevada and Wisconsin.
-Though Harris has predicted to have a majority in Ohio and Florida, the disappointing results in the other states makes her lead in these states less signifiant.
-Moreover, looking at the general NA category,representing all non-state specific polls, Harris' predicted approval rating hovers at 48%, which means she is expected to lose the popular vote to Trump.
+In our model, by separating state-specific polls, we can analyze and model the poll results by states where state-specific polls were conducted.
+Examining @fig-state-comparison, common battleground states such as Arizona, Pennsylvania, Wisconsin, and Michigan are polled more frequently at the state level than partisan states such as California and New York.
+Examining the results at these key battleground states, the results show that Harris' popularity recently fell below 50% during October and November in key states such as Arizona, Michigan, Pennsylvania, Nevada, and Wisconsin.
+Though Harris is predicted to have a majority in Ohio and Florida, the disappointing results in the other states make her lead in these states less powerful.
 
-@fig-state-summary shows that Harris is projected to have a minor lead with 50.4% of the popular vote in five key states: Virginia, Ohio, New Hampshire, Minnesota, and Florida.
-However, in the other states, she is expected to lose the popular vote.
-Pennsylvania, Nevada, Michigan, and Georgia are critical states for determining the outcome of the election.
-As Harris has a noticeable projected minority in these states, she not only is expected to lose the popular vote, but Trump has a projected lead over Harris in these battleground states, hurting her chances at being elected.
-
-Thus, our model predicts that Harris is unlikely to win the general popular vote or key swing states. Consequently, we anticipate Donald Trump will win the 2024 U.S. Presidential election.
+@fig-state-summary shows Harris with a slight lead of 50.4% in the popular vote across five key states: Virginia, Ohio, New Hampshire, Minnesota, and Florida.
+In other states, she is expected to either be in a close race or lose the popular vote.
+Pennsylvania, Nevada, Michigan, and Georgia are crucial battlegrounds, where Harris's minority share and Trump’s projected lead reduce her chances of winning the election.
+Thus, our model predicts that Harris is unlikely to win the general popular vote or key swing states.
+Consequently, we anticipate Donald Trump will win the 2024 U.S.
+Presidential election.
 
@@
 ```{r}
-#| label: tbl-state-missing
-#| fig-cap: Summary of missing and total dates across key states
 #| echo: false
-#| tbl-align: center
+#| message: false
+#| warning: false
+#| label: fig-state-comparison
+#| fig-cap: State-by-State comparison of predicted percentages over time
+#| fig-align: center
 
 state_data <- prediction %>% filter(national_poll == 0)
+
 # Identify missing data
@@
 
-# Display the summary statistics in a nicely formatted table
-missing_data_summary %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-```{r}
-#| label: fig-state-comparison
-#| fig-cap: State-by-State comparison of predicted percentages over time
-#| echo: false
-#| fig-align: center
-
 # Plot with points for sparse data
-ggplot(prediction, aes(x = end_date, y = predicted_pct, color = state)) +
+ggplot(state_data, aes(x = end_date, y = predicted_pct, color = state, group = state)) +
   geom_line(na.rm = TRUE) +
@@
 #| fig-align: center
-
 # Filter to keep only the most recent entry for each state based on 'end_date'
@@
 
-If our prediction is correct and Trump returns to office, we would see a shift in the United States' domestic and foreign policy compared to the Biden administration.
-Donald Trump's past presidency from 2016 to 2020 was marked with tax cuts, stronger border enforcement, isolationist foreign policy, and a massive rollback on environmental regulations [@BBCNews2024].
+If our prediction is correct and Trump returns to office, we will see a shift in the United States' domestic and foreign policy compared to the Biden administration.
+Donald Trump's presidency from 2016 to 2020 was marked by tax cuts, stronger border enforcement, isolationist foreign policy, and a massive rollback on environmental regulations [@BBCNews2024].
 During this campaign cycle, Trump has largely promoted similar policies, representing a stark contrast to Harris' campaign.
-The implication of the paper is that if Trump wins the election as projected, then we would see the re-introduction of these policies and a different America on the global stage.
+The paper implies that if Trump wins the election as projected, then we will see the re-introduction of these policies and a different America on the global stage.
 As the United States election is one of the most important international events of this year, predicting the event's outcome is critical for foreign governments and businesses to guide strategy on US relations and regulations.
 
-Moreover, the paper has significant lessons for the Harris campaign.
+Moreover, the paper has lessons for the Harris campaign.
 Harris should prioritize swing states where small shifts in public opinion can determine electoral outcomes.
-Harris' campagin may need to adjust its messaging and outreach in battleground states to mitigate her projected losses.
-Also, shown through @fig-predicted-and-real, popularity can flucturate based on public enthusiasm and external factors.
-Campagins must focus on mobilizing their voter base effectively in the weeks leading up to the election.
+Harris' campaign may need to adjust its messaging and outreach in battleground states to mitigate her projected losses.
+Also, as shown through @fig-predicted-and-real, popularity can fluctuate based on public enthusiasm and external factors.
+Campaigns must focus on mobilizing their voter base effectively in the weeks leading up to the election.
 Though the paper can help guide the Harris campaign leading up to the election, it also can serve to help future candidates in guiding their strategy surrounding shaping public opinion and prioritizing swing states.
@@
 
-The first weakness in our study is the discrepancy of datapoints between states.
+The first weakness in our study is the discrepancy of data points between states.
 @fig-state-summary shows that certain states have more frequent state-specific polls than others.
-While states like California, Montana, and New York have historically leaned towards one party than the other, the discrepancy in datapoints between states is a weakness in our data and hurts our abilty to draw conclusions from our data, as it is difficult to make out trends.
-Therefore, for future studies, a next potential step could be to conduct more state-specific polls and map out a candidate's popularity across each state.
+While states like California, Montana, and New York have historically leaned towards one party than the other, the discrepancy in data points between states is a weakness in our data and hurts our ability to conclude our data, as it is difficult to make out trends.
+Therefore, for future studies, the next potential step could be to conduct more state-specific polls and map out a candidate's popularity across each state.
 
@@
 While the popular vote is a valuable metric, the Electoral College system means that a candidate can lose the popular vote and still win the presidency, as seen in past elections.
-Harris' projected loss in key states, combine dwith a predicted minority in the national polls, signals that simply gaining a majority in the popular vote may not be sufficient for electoral success under the current system.
-Therefore, the next step for this study is to further incorporate the Electoral College and its methodology to the study, requiring more state-specific polling and adjusting the estimand.
+Harris' projected loss in key states, combined with a predicted minority in the national polls, signals that simply gaining a majority in the popular vote may not be sufficient for electoral success under the current system.
+Therefore, the next step for this study is to further incorporate the Electoral College and its methodology into the study, requiring more state-specific polling and adjusting the estimand.
 
 Finally, as we are working with polling data, there is potential for bias.
-As mentioned earlier in @sec-data-pollrecency, potential for bias may disproportionately influence model predictions.
+As mentioned earlier in @sec-data-pollrecency, the potential for bias may disproportionately influence model predictions.
 This has broader implications for the accuracy of electoral forecasts and highlights the need for continually updated polling to provide a more accurate reflection of voter sentiment as Election Day approaches.
@@
 
-YouGov employs a panel-based methodology for conducting online surveys, recruiting participants who voluntarily join their platform.
-Their surveys span topics like politics, policy opinions, and voter behavior.
-In the context of election polling, YouGov applies Multilevel Regression and Post-Stratification (MRP) models to predict outcomes by blending respondent data with external data like voter files [@yougov_mrp_2024].
+YouGov utilizes a panel-based methodology for conducting online surveys and recruiting participants who voluntarily join their platform.
+Survey topics range widely, covering areas such as political views, policy opinions, and voter behavior.
+For election polling, YouGov employs a Multilevel Regression and Post-stratification (MRP) model that integrates respondent data with external sources like voter files to enhance the accuracy of outcome predictions [@yougov_mrp_2024].
 
-The target population for YouGov’s surveys typically consists of U.S.
-adults, with specific samples drawn for each survey depending on the topic.
-YouGov’s sample frame consists of individuals who have opted into their online panel.
-To ensure representativeness, YouGov sets demographic quotas, stratifying by variables such as age, gender, and political affiliation.
-Additionally, YouGov monitors and adjusts their sample post-survey using MRP to correct for demographic imbalances and potential sampling bias.
+Surveys target U.S.
+adults, with samples carefully selected based on survey topics.
+The sampling process begins with YouGov’s online panel, comprised of individuals who join voluntarily.
+To improve representativeness, YouGov establishes demographic quotas by age, gender, and political affiliation and applies MRP-based post-survey adjustments.
+These steps help to mitigate demographic imbalances and minimize sampling bias [@yougov_mrp_2024].
 
-YouGov’s panel is recruited through various online methods, including digital advertisements and direct outreach.
-This recruitment strategy leads to a convenience sample, meaning participants self-select to join the panel, which can introduce selection bias.
-Although YouGov attempts to offset this bias using quotas and post-stratification weighting, the sample still primarily consists of individuals comfortable with online surveys.
-This method may under-represent certain demographics, such as older adults or individuals with limited internet access.
+While YouGov’s non-probability sampling method offers a fast and cost-effective approach to data collection, it may introduce self-selection bias.
+Though MRP adjustments use broad population data to reduce these biases, reliance on a non-random sample can still result in skewed outcomes.
+Survey format and design also play roles in influencing participant responses, potentially impacting representativeness across groups with differing levels of digital literacy or social trust [@yougov_poll_methodology].
 
-YouGov’s use of non-probability sampling—drawing respondents from a convenience sample—provides cost efficiency and quick data collection.
-However, the trade-off lies in the introduction of potential biases due to self-selection.
-While YouGov’s MRP model corrects some biases by leveraging population-level data for small samples, the non-random nature of the sample may still skew results.
-For instance, respondents who engage in online surveys may systematically differ from those who avoid them, such as individuals with lower levels of social trust.
+Non-response bias presents additional challenges, as individuals who abstain from participation may differ in ways from those who complete surveys.
+YouGov addresses this issue by applying post-survey weighting to align responses with population characteristics and by using imputation to estimate missing data based on observed patterns[@yougov_panel_methodology].
+Despite these steps, non-response bias can still affect predictions, particularly when groups less inclined to participate, such as those with lower levels of education or trust, are underrepresented [@cloudresearch_nonresponse_bias].
 
-Non-response bias is one of the main concerns in YouGov’s methodology.
-Individuals who opt out of surveys may systematically differ from those who participate.
-YouGov employs several strategies to mitigate this issue, including post-survey weighting, which adjusts the survey data to reflect population characteristics, and imputation techniques, which estimate missing responses based on patterns in the available data.
-Despite these efforts, non-response bias can still affect the accuracy of predictions, particularly when certain demographic groups (e.g., less-educated or low-trust individuals) are less likely to participate.
-
-YouGov focuses on clear and concise survey design, reducing participant fatigue and improving data quality.
-Surveys typically avoid overly complex or leading questions, which can skew the results.
-However, question-wording remains an important area, as even subtle changes in expression can affect responses.
-For example, YouGov's research shows how question-wording can affect voter perceptions and reported support for policies.
-Furthermore, prolonged surveys can lead to increased dropout rates, potentially resulting in non-response bias.
-To counteract this, YouGov often dynamically adjusts questions based on previous responses to keep participants engaged by incorporating survey logic.
-
-YouGov’s panel-based methodology, combined with statistical techniques like MRP, offers a robust approach to polling.
-Their methods effectively address many challenges associated with non-random sampling and non-response bias.
-However, limitations persist, particularly regarding the biases inherent in self-selection and question framing.
-Overall, while YouGov’s methodology is well-suited to large-scale surveys, particularly in politics, careful consideration of their sampling and non-response strategies is necessary to fully interpret their findings.
+YouGov’s combination of a panel-based approach and MRP modeling provides a strong framework to manage non-random sampling challenges and mitigate non-response bias.
+However, limitations persist, particularly regarding self-selection and question phrasing.
+While well-suited for large-scale political surveys, this methodology requires careful interpretation of sampling and non-response strategies [@yougov_mrp_2024].
 
@@
 Because the U.S.
-operates on the Electoral College system, where state-affiliation matters, I would use Stratified Random Sampling to ensure each state is represented proportionally in the survey, improving accuracy and representativeness.
+operates on the Electoral College system, where state affiliation matters, we would use Stratified Random Sampling to ensure each state is represented proportionally in the survey, improving accuracy and representativeness.
 My target population would be voter-eligible citizens in the United States who are above 18 years old, with a sampling frame utilizing voter registration databases stratified by key demographics, including age, race, gender, geographic region, income, and education level.
-I aim for a sample size of around 10,000 respondents to achieve a reasonable margin of error, ensuring robust findings, though I expect the actual sample may be less due to various factors.
+We aim for a sample size of around 10,000 respondents to achieve a reasonable margin of error, ensuring findings, though we expect the actual sample may be less due to various factors.
 
-To recruit respondents, I would primarily utilize online methods, such as surveys on social media platforms like Facebook and Instagram to engage younger audiences, and interactive voice response systems for older or less tech-savvy individuals.
-To ensure the data collected is valid, I will cross-check responses with voter registration databases and implement logic checks within the survey to detect inconsistencies.
+To recruit respondents, we would primarily utilize online methods, such as surveys on social media platforms like Facebook and Instagram to engage younger audiences, and interactive voice response systems for older or less tech-savvy individuals.
+To ensure the data collected is valid, we will cross-check responses with voter registration databases and implement logic checks within the survey to detect inconsistencies.
 For instance, if a respondent claims to have already voted but indicates they’re unlikely to vote, this response will be flagged as unreliable.
 
-Addressing non-response bias is crucial, and I plan to over-sample underrepresented groups while employing multiple attempts to contact individuals online and conducting weekly polls.
+Addressing non-response bias is important, and we plan to over-sample underrepresented groups while employing multiple attempts to contact individuals online and conducting weekly polls.
 Frequent surveys will allow me to aggregate results using a moving average to smooth out short-term fluctuations in responses.
-I will also implement checks to prevent duplicate entries in the online survey and track phone responses to ensure unique participants.
+we will also implement checks to prevent duplicate entries in the online survey and track phone responses to ensure unique participants.
 Given that the U.S.
-does not conduct elections based on a popular vote, I will weight the survey results to reflect the voting population as indicated by U.S.
+does not conduct elections based on a popular vote, we will weight the survey results to reflect the voting population as indicated by U.S.
 Census data and voter turnout estimates, considering demographic factors such as race, gender, and age.
 
-To allocate the \$100,000 budget, I will spend \$20,000 on a survey platform subscription, \$50,000 on targeted digital ads and phone surveys for recruitment, \$20,000 on data analysts and poll aggregation services, and \$10,000 for re-contact surveys and miscellaneous costs.
-However, I recognize that this budget may be on the lower end to ensure a high-quality poll and that the actual costs might exceed this amount.
-Most expenses will arise from data collection, and I have accounted for this in my budget planning.
+To allocate the \$100,000 budget, we will spend \$20,000 on a survey platform subscription, \$50,000 on targeted digital ads and phone surveys for recruitment, \$20,000 on data analysts and poll aggregation services, and \$10,000 for re-contact surveys and miscellaneous costs.
+However, we recognize that this budget may be on the lower end to ensure a high-quality poll and that the actual costs might exceed this amount.
+Most expenses will arise from data collection, and we have accounted for this in our budget planning.
 
@@
 @fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
-Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since he trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.1, further supporting convergence.
+Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since the trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.1, further supporting convergence.
 

### 2024-11-04 paper/paper.qmd

@@
 #| echo: false
-#| message: false
-#| warning: false
-#| label: fig-populationtype
-#| fig-cap: Distribution of population type
 #| fig-width: 5
+#| fig-height: 5
 #| fig-align: center
@@
 ```{r}
+#| label: tbl-state-missing
+#| tbl-cap: Summary of missing and total dates across key states
 #| echo: false
@@
             total_dates = n())
+
+# Display the summary statistics in a nicely formatted table
+missing_data_summary %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-state-comparison
+#| fig-cap: State-by-State comparison of predicted percentages over time
+#| echo: false
+#| message: false
+#| fig-align: center
 

### 2024-11-04 paper/paper.qmd
paper update

@@
 #| echo: false
+#| message: false
+#| warning: false
+#| label: fig-populationtype
+#| fig-cap: Distribution of population type
 #| fig-width: 5
@@
 ```{r}
-#| label: tbl-state-missing
-#| tbl-cap: Summary of missing and total dates across key states
 #| echo: false
@@
 #| fig-align: center
-
+#| 
 state_data <- prediction %>% filter(national_poll == 0)
@@
             total_dates = n())
-
-# Display the summary statistics in a nicely formatted table
-missing_data_summary %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-```{r}
-#| label: fig-state-comparison
-#| fig-cap: State-by-State comparison of predicted percentages over time
-#| echo: false
-#| message: false
-#| fig-align: center
 
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-state-summary
 #| fig-cap: Summary of Kamala Harris' predicted popular vote by key states
-#| echo: false
-#| message: false
 #| fig-align: center

### 2024-11-04 paper/paper.qmd
paper update - modified graph heights

@@
 Despite steady support at 48%, Harris is projected to lose the national popular vote and key battleground states to Donald Trump.
-Our findings emphasize the significance of polling methodologies and their role in predicting U.S.
+Our findings highlight the impact of polling methodologies and their role in shaping predictions of U.S.
 policy outcomes.
@@
 ```
-\newpage
 
@@
 
-\newpage
-
 ### Population Type {#sec-data-populationtype}
@@
 ```
-\newpage
+
 ```{r}
@@
 #| fig-width: 5
-#| fig-height: 5
+#| fig-height: 2.5
 #| fig-align: center
@@
 #| fig-width: 5 
-#| fig-height: 5
+#| fig-height: 2.5
 #| fig-align: center
@@
 ```
-
-\newpage
 
@@
 
-\newpage
-
 ```{r}
@@
 Please find the proposed question list for the online survey attached [here](https://forms.gle/AFqodUEk48o864aN6).
+\newpage
 
